---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting Started with R at Reed

R is a versatile tool, and at Reed R is used across departments and in various ways. Our goal is to help you navigate this tool in any setting that you might use it. This document will be an introduction to the Reed RStudio Server, Desktop R, and the R thesis template. 

### Reed RStudio Server

There are a couple of great advantages to using R on the Reed RStudio server:

- Accessibility from any computer
- Avoid possible messy installation
- Provides some storage space for files, data, and images

To access the server, head to [rstudio.reed.edu](https://rstudio.reed.edu) and log in using your Kerberos credentials.

#### Installing Libraries

Most of the libraries that you will need are already installed on the server, but if you want to use a package that is not already on the server you can use `install.package("package_name")`. If you run into errors resulting from installing such packges, contact Kbott (kbott@reed.edu).

#### Jointly Editing Documents

We strongly recommend having one active editor for a shared project. (While you may be used to the collaborative editing/versioning functions in Google Docs, that functionality does not transfer to projects on the RStudio server.) If the server detects that multiple editors are accessing the file, it will block further editing of that file. If the server gives you a saving error, copy/paste your code into a text editor or Google Doc, since the server will not save your work.

If you're looking into jointly editing documents you can look into using [git and github] (https://happygitwithr.com/rstudio-git-github.html).

#### The server is broken / down / having problems

Every five minutes, the RStudio Server checks to make sure it is running smoothly; if there is a glitch, the server will restart. This restart should solve any server-level problems.

#### Error occurred during transmission (broken sessions)

At some point, you may encounter an error when you log in: RStudio initialization error, ‘Error occurred during transmission’. Sometimes RStudio user sessions encounter errors of some stripe; while this might be annoying, it is also a problem you can solve. To exit a misbehaving session or create a new session, visit (https://rstudio.reed.edu/home).

### Desktop R

If you are using R for projects outside of specific class assignments that are accessed and submitted using the Reed RStudio server, and especiallly if you are using sensitive data, you will want to install the Desktop version of RStudio.

There are two things you need to download in order to use R on your computer: R and RStudio. R is the coding language, and RStudio is the Integrated Development Environment (IDE), which makes R easy to use.

#### On MacOSX

You can download and install R for MacOSX at the [CRAN downloads site](https://cran.r-project.org/), where you'll follow the link to "Download R for (Mac) OS X". Navigate to the appropriate release for your computer's OS. To find your computer's OS, click the top left apple icon, and click "About this Mac". You will see both a name (e.g. "High Sierra", "Catalina", or "El Capitan") and a number ("Version __.__._"). Back on the download site, each download will have a description of what OS it requires in order to run. Choose the newest version possible for your OS (your OS should be equal to or higher than the one listed in the description of the download). Click on the link next to the description to download R. Once the file has downloaded, click it to proceed to installation, leaving all default settings as is.

Next we need to install RStudio. To do this, navigate to the [RStudio downloads page](https://rstudio.com/products/rstudio/download/), and scroll down to "All Installers". Choose the download that matches your OS, most likely "MacOS 10.13+". Click the download link, open the file when it has downloaded, and install with the default settings as you did with R. 

Now you should be able to open RStudio and begin working! Remember to install packages as you need them, either using `install.packages()`, or by navigating to the "Packages" window within RStudio, where you can search for and install packages. You only need to install packages once in order to load and use them in the future.

#### On Windows

You can download and install R for Windows at the [CRAN downloads site](https://cran.r-project.org/), where you'll follow the link to "Download R for Windows". From here click "base", and then "Download R 4.0.3 for Windows". Once the file has downloaded, click it to proceed to installation, leaving all default settings as is.

Next we need to install RStudio. To do this, navigate to the [RStudio downloads page](https://rstudio.com/products/rstudio/download/), and scroll down to "All Installers". Choose the download for Windows ("Windows 10/8/7"). Click the download link, open the file when it has downloaded, and install with the default settings as you did with R. 

Now you should be able to open RStudio and begin working! Remember to install packages as you need them, either using `install.packages()`, or by navigating to the "Packages" window within RStudio, where you can search for and install packages. You only need to install packages once in order to load and use them in the future.

### Writing Your Thesis in `thesisdown`

The `thesisdown` package provides an R Markdown template for writing your senior thesis, which is particularly helpful for easily integrating R code and graphics into your thesis manuscript. Instructions for how to install this package and use the template can be found in the [GitHub repository](https://github.com/ismayc/thesisdown).

#### Uploading Data to the Reed Rstudio Server

You might have some data on your computer that you'd like to work with on the server. You can upload data from your computer by:

* logging in to the [RStudio server](https://rstudio.reed.edu) and opening up a session
* find the "Files" pane in the bottom right
* navigate to the folder that you'd eventually like your file to be in
* click the "Upload" button (with a yellow, up-facing arrow beside it)
* choose the file on your computer

The data is now uploaded to the server! Now, you'll read in the data to R in the same way you would if you were working with data stored on your computer with the desktop version of R. See our [loading data](LINK-HERE) page for more information.

### Meet the Palmer Penguins

Throughout these tutorials, we'll be making use of a dataset called `penguins` from the `palmerpenguins` package. The dataset contains various different measurements for penguins of three species on three islands in the Palmer Archipelago, Antarctica.

```{r, echo = FALSE}
url <- "https://github.com/allisonhorst/palmerpenguins/raw/master/man/figures/lter_penguins.png"

knitr::include_graphics(url)
```

If you want to work along with any of the examples on this website and need to load the data, you will first need to install the palmerpenguins package:

```{r, eval = FALSE}
install.packages("palmerpenguins")
```

Note that you only need to _install_ the package once. Once you have installed the package, though, you will need to _load_ the package every time you want to use the data. You can do so with the following code:

```{r}
library(palmerpenguins)
```

Finally, loading the data with the `data()` function:

```{r}
data("penguins")
```

You might notice that the dataset does not seem to be fully loaded in—beside `penguins`, there is probably a `<Promise>` symbol. To _really_ load the data in, click on `<Promise>`. `penguins` is now ready to go.

Let's check `penguins` out:

```{r}
penguins
```

In this data, every row is a unique penguin. For each penguin, we have the `species` of the penguin, which `island` it lives on, its `sex`, the `year` the measurements were taken, and some body size measurements.

To learn more about all that we can do with this data, check out the remaining tutorials.
