---
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting Started with R at Reed

R is a versatile tool, and at Reed R is used across a number of departments and disciplines. Our goal is to help you navigate this tool in any setting that you might use it. This document will be an introduction to the Reed RStudio Server, Desktop R, and the R thesis template. 

---------------------------------------------------------
### Reed RStudio server

There are a couple of advantages to using R on the Reed RStudio Server:

- Accessible from any computer with an internet connection
- Avoids the need to install software
- Provides some storage space for files, data, and images

To access the server, head to [rstudio.reed.edu](https://rstudio.reed.edu) and log in using your Kerberos credentials.

#### Installing libraries

Most of the libraries that you will need are already installed on the server and globally available. If you want to use a package that is not already on the server you can use `install.package("packagename")` to install a library within your own user space on the server. If you run into errors installing or using packages, contact our RStudio administrators at [rstudio-support@reed.edu](mailto:rstudio-support@reed.edu).

#### Uploading data to the Reed Rstudio server

You might have some data on your computer that you would like to work with on the server. To upload data from your computer:

* log in to the [RStudio server](https://rstudio.reed.edu) to open an RStudio session
* find the "Files" pane in the bottom right
* navigate to the folder (on the server) where you want to store your data
* click the "Upload" button (with a yellow, up-facing arrow beside it)
* choose the file on your computer and click "Open"

The data is now uploaded to the server. You can now load the data into R in the same way you would if you were working with data stored on your computer with the desktop version of R. See our [loading data](LINK-HERE) page for more information.

#### Jointly editing documents

One of the advantages to working on RStudio server is the ability to easily share projects with collaborators. When working on shared projects, we strongly recommend having **one** active editor for a project, because RStudio Server does not have the same simultaneous editing ability that a tool like Google Docs does. If the server detects that multiple editors are accessing the file, it will temporarily block further editing of that file and may not save your work. If the server gives you a saving error, copy and paste your code into a text editor or Google Doc. You can then close the file, wait for your partner to be done with their edits, and then reopen the file to implement your changes.

If you are want to jointly edit documents, you can look into using [Git and GitHub] (https://happygitwithr.com/rstudio-git-github.html).

#### The server is broken / down / having problems

Every five minutes, the RStudio Server checks to make sure it is running smoothly; if there is a glitch, the server will restart. This restart should solve any server-level problems.

#### Error occurred during transmission (broken sessions)

At some point, you may encounter an error when you log in: RStudio initialization error, ‘Error occurred during transmission’. Sometimes RStudio user sessions encounter errors of some stripe; while this might be annoying, it is also a problem you can solve. To exit a misbehaving session or create a new session, visit [https://rstudio.reed.edu/home](https://rstudio.reed.edu/home).

---------------------------------------------------------

### Desktop R

If you are using R for projects outside of specific class assignments that are accessed and submitted using the Reed RStudio Server, if you are using sensitive data, or if you want to be able to work in R independent of an internet connection, you will want to install the desktop version of R and RStudio.

There are two things you need to download in order to use R on your computer: R and RStudio. R is the coding language, and RStudio is the Integrated Development Environment (IDE) that makes R easy to use.

#### Downloading + installing for MacOSX

You can download and install R for MacOSX at the [CRAN downloads site](https://cran.r-project.org/), where you will follow the link to "Download R for (Mac) OS X". Navigate to the appropriate release for your computer's OS. To find your computer's OS, click the top left Apple icon, and click "About this Mac". You will see both a name (e.g. "High Sierra", "Catalina", or "El Capitan") and a number ("Version __.__._"). Back on the R download site, each download will have a description of what OS it requires in order to run. Choose the newest version possible for your OS (your OS should be equal to or higher than the one listed in the description). Click on the link next to the description to download R. Once the file has downloaded, click it to proceed to installation, leaving all default settings as they are.

Next you need to install RStudio. To do this, navigate to the [RStudio downloads page](https://rstudio.com/products/rstudio/download/) and scroll down to "All Installers". Choose the download that matches your OS, most likely "MacOS 10.13+". Click the download link, open the file when it has downloaded, and install with the default settings as you did with R.

With both R and RStudio installed, you should be able to open RStudio and begin working. Remember to install packages as you need them, either using `install.packages("packagename")`, or by navigating to the "Packages" window within RStudio, where you can search for and install packages. You only need to install a package once in order to load (`library(packagename)`) and use it in the future, but you need to load a package every time you use it.

#### Downloading + installing for Windows

You can download and install R for Windows at the [CRAN downloads site](https://cran.r-project.org/), where you'll follow the link to "Download R for Windows". From here click "base", and then "Download R 4.0.3 for Windows". Once the file has downloaded, click it to proceed to installation, leaving all default settings as they are.

Next you need to install RStudio. To do this, navigate to the [RStudio downloads page](https://rstudio.com/products/rstudio/download/) and scroll down to "All Installers". Choose the download for Windows ("Windows 10/8/7"). Click the download link, open the file when it has downloaded, and install with the default settings as you did with R. 

Now you should be able to open RStudio and begin working. Remember to install packages as you need them, either using `install.packages("packagename")`, or by navigating to the "Packages" window within RStudio, where you can search for and install packages. You only need to install a package once in order to load (`library(packagename)`) and use it in the future, but you need to load a package every time you use it.

---------------------------------------------------------

### Writing your thesis in `thesisdown`

The `thesisdown` package provides an R Markdown template for writing your senior thesis. This is particularly helpful for easily integrating R code and graphics into your thesis manuscript. Instructions for how to install this package and use the template can be found in the [GitHub repository](https://github.com/ismayc/thesisdown).

---------------------------------------------------------

### Meet the Palmer Penguins

Throughout these tutorials, we will be making use of a dataset called `penguins` from the `palmerpenguins` package. The dataset contains various different measurements for penguins of three species on three islands in the Palmer Archipelago, Antarctica.

```{r, echo = FALSE}
url <- "https://github.com/allisonhorst/palmerpenguins/raw/master/man/figures/lter_penguins.png"

knitr::include_graphics(url)
```

If you want to work along with any of the examples on this website and need to load the data, you will first need to install the palmerpenguins package:

```{r, eval = FALSE}
install.packages("palmerpenguins")
```

Note that you only need to _install_ the package once. Once you have installed the package, though, you will need to _load_ the package every time you want to use the data. To load the package:

```{r}
library(palmerpenguins)
```

Now you can load the data with the `data()` function:

```{r}
data("penguins")
```

You might notice that the dataset does not seem to be fully loaded in. Beside `penguins`, there is probably a `<Promise>` symbol. To _really_ load the data in, click on `<Promise>`. `penguins` is now ready to go.

We can now explore `penguins`:

```{r}
penguins
```

In this data, every row is a unique penguin. For each penguin, we have the `species` of the penguin, which `island` it lives on, its `sex`, the `year` the measurements were taken, and some body size measurements.

To learn more about all that we can do with this data, check out the remaining tutorials.
