---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(knitr)
```

# Wrangling Data

_explanation/something goofy about data wrangling_

### Data Frames and `tibble`s

_yada yada about data frames and tibbles_ - biggest thing to note here is that the print method (i.e. what happens when you just type the name of a data frame) is much better than that for normal data frames

### Tidy Data Principles, Reshaping Data, and `tidyr`

Packages in the `tidyverse`, like `ggplot2` and `dplyr`, are built to work with tidy data. The `tidyr` package can help us tidy up a dataset by rearranging how data is structured.

Briefly, for a dataset to be tidy, it must follow these three rules:

1. Each variable must have its own column

2. Each observation must have its own row

3. Each value must have its own cell

Let's take a look at a summarized version of the `penguins` data as an example:

```{r, echo = FALSE, message = FALSE, warning = FALSE}
penguins_sum <- penguins %>%
  group_by(island, year) %>%
  summarize(mean_body_mass_g = mean(body_mass_g, na.rm= TRUE))

penguins_sum
```



Each variable has its own column, each observation (the sample of penguins on a given island in a given year) is a row, and each value has its own cell.

This dataset, though, could also look like this:

```{r}
penguins_sum %>%
  pivot_wider(id_cols = c("island", "year"), names_from = year, values_from = mean_body_mass_g)
```

This dataset technically contains the same data, but in this version, the mean body mass is held in three different columns!

Sometimes, the data you're asked to analyze might look more like the second, untidy version than the first, tidy version. How do you move from one to the other? Introducing `tidyr`!

Let's start off with the first version:

```{r, message = FALSE, warning = FALSE}
penguins_sum
```

We want to "pivot" this data from a "long" format to a "wide" format. The `pivot_wider()` function from `tidyr` will help us to do that:

```{r, warning = FALSE, message = FALSE}
penguins_wide <- penguins_sum %>%
  pivot_wider(id_cols = c("island", "year"), 
              names_from = year, 
              values_from = mean_body_mass_g)

penguins_wide
```

The `id_cols` (read: ID columns) are the variables that, together, identify what makes a row unique in the original data, `names_from` determines what columns in the old data will be the names of the new columns, and `values_from` is the name of the variable in the old data that will be the information contained inside the new columns.

What if we want pivot from wider to longer, though? You might have guessed it! `pivot_longer()`! Starting with the widened version of the data:

```{r}
penguins_wide %>%
  pivot_longer(cols = c("2007", "2008", "2009"), 
               names_to = "year",
               values_to = "mean_body_mass_g")
```

As you might have noticed, `pivot_longer()` and `pivot_wider()` are inverse operations! Pivoting a widened dataset to a longer format gives you back the original dataset, and vice versa.

If you're interested in learning more about tidy data and pivoting, we recommend the chapter in [R for Data Science](https://r4ds.had.co.nz/tidy-data.html)!

### Transforming Data With `dplyr`

_yada yada about data transformation_