---
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting Started with R at Reed

### Reed RStudio server

There are a couple of advantages to using R on the Reed RStudio Server:

- Accessible from any computer with an internet connection
- No need to install software
- Provides some storage for files, data, and images
- Useful for collaborative work (caution: see notes below)

To access the server, head to [rstudio.reed.edu](https://rstudio.reed.edu) and log in using your Kerberos credentials.

#### Installing libraries

Most of the libraries that you will need are already installed on the server. If you want to use a package that is not already on the server you can type `install.package("packagename")` in the console; this will install a library within your own user space on the server. If you run into errors installing or using packages, contact the RStudio server administrators at [rstudio-support@reed.edu](mailto:rstudio-support@reed.edu).

#### Uploading data to the Reed Rstudio server

You might have some data on your computer that you would like to work with on the server. To upload data from your computer:

* log in to the [RStudio server](https://rstudio.reed.edu) to open an RStudio session
* in the bottom right corner of your screen, locate the "Files" pane
* within the "Files" pane, navigate to the folder (on the server) where you want to store your data
* click the "Upload" button (with a yellow, up-facing arrow beside it)
* choose the file on your computer and click "Open"

The data is now uploaded to the server. You can now load the data into R in the same way you would if you were working with data stored on your computer with the desktop version of R. See our [loading data](https://www.reed.edu/data-at-reed/resources/R/index.html) section for more information.

#### Jointly editing documents

One of the advantages to working on RStudio server is the ability to easily share projects with collaborators. When working on shared projects, it is best to have **one** active editor for a project, because RStudio Server does not have the simultaneous editing ability that a tool like Google Docs does. If the server detects that multiple editors are accessing the file, it will temporarily block further editing of that file and may not save your work. If the server gives you a saving error, copy and paste your code into a text editor or Google Doc to retain your work. You can then close the file on the Rstudio server, wait for your partner to be done with their edits, and then reopen the file to implement your changes (copy/paste the text you had saved).

If you want to jointly edit documents with collaborators, you can look into using [Git and GitHub](https://happygitwithr.com/rstudio-git-github.html).

#### Help! The server is broken / down / having problems

Every five minutes, the RStudio Server checks to make sure it is running smoothly; if there is a glitch, the server will restart. This restart should solve any server-level problems.

#### Error occurred during transmission (broken sessions)

At some point, you may encounter an error when you log in: RStudio initialization error, ‘Error occurred during transmission’. Sometimes RStudio user sessions encounter errors of some stripe; while this might be annoying, it is also a problem you can solve. To exit a misbehaving session or create a new session, visit [https://rstudio.reed.edu/home](https://rstudio.reed.edu/home).

---

### Desktop R

You may wish to install and use the desktop version of R and RStudio if any of the following are true:

- you are using R for projects outside of specific class assignments that are accessed and submitted using the Reed RStudio Server
- you are using sensitive data
- you want to be able to work in R independent of an internet connection

There are two things you need to download in order to use R on your computer: R and RStudio. R is underlying program and language, and RStudio is the Integrated Development Environment (IDE) that makes R more friendly to use.

#### Downloading + installing for MacOSX

You can download and install R for MacOSX at the [CRAN downloads site](https://cran.r-project.org/), where you will follow the link to "Download R for (Mac) OS X". Navigate to the appropriate release for your computer's operating system (OS). To find your computer's OS, click the top left Apple icon, and click "About this Mac". You will see both a name (e.g. "Catalina", "El Capitan", or "Big Sur") and a number ("Version __.__._"). 

Back on the R download site, each download will have a description of what OS it requires in order to run. Choose the newest version possible for your OS (your OS should be equal to or higher than the one listed in the description). Click on the link next to the description to download R. Once the file has downloaded, click it to proceed to installation, leaving all default settings as they are.

Next you need to install RStudio. To do this, navigate to the [RStudio downloads page](https://rstudio.com/products/rstudio/download/) and scroll down to "All Installers". Choose the download that matches your OS, most likely "MacOS 10.13+". Click the download link, open the file when it has downloaded, and install with the default settings as you did with R.

With both R and RStudio installed, you should be able to open RStudio and begin working. Remember to install packages as you need them, either using `install.packages("packagename")`, or by navigating to the "Packages" window within RStudio, where you can search for and install packages. You only need to install a package once in order to load (`library(packagename)`) and use it in the future. Remember that you need to load a package every time you use it.

#### Downloading + installing for Windows

You can download and install R for Windows at the [CRAN downloads site](https://cran.r-project.org/), where you'll follow the link to "Download R for Windows". From here click "base", and then "Download R [version number] for Windows". Once the file has downloaded, click it to proceed to installation, leaving all default settings as they are.

Next you need to install RStudio. To do this, navigate to the [RStudio downloads page](https://rstudio.com/products/rstudio/download/) and scroll down to "All Installers". Choose the download for Windows ("Windows 10/8/7"). Click the download link, open the file when it has downloaded, and install with the default settings as you did with R. 

Now you should be able to open RStudio and begin working. Remember to install packages as you need them, either using `install.packages("packagename")`, or by navigating to the "Packages" window within RStudio, where you can search for and install packages. You only need to install a package once in order to load (`library(packagename)`) and use it in the future. Remember that you need to load a package every time you use it.

---

### Writing up your results

#### R Script

R script files (`.R`) are a good starting point for learning to code in R, with simple formatting and the ability to integrate comments with your code. You can easily alternate between code and regular text using comments with a pound sign at the beginning of a line that contains text (`#like this`). Comments are useful for a number of reasons, including (1) to remind yourself what your code is doing (or denote problems/questions as you work), (2) to explain your code to others, and (3) to provide general organization and clarification throughout your code.

To learn more about R Scripts, visit the [Workflow: scripts](https://r4ds.had.co.nz/workflow-scripts.html) chapter in [R for Data Science](https://r4ds.had.co.nz/).

#### R Markdown

R Markdown (`.Rmd`) filed provide a way to present final results with R. Markdown documents can integrate text, code, and output; this combination provides flexibility for presenting your work. R Markdown files `Knit` into output documents, which are most commonly `.pdf` (for assignments and "white papers") and `.html` (for web content). When working in a Markdown file in RStudio, you can find the `Knit` button in the tool bar near the top of the file. 
Note: One caveat for R Markdown is that sometimes knitting your document can bring up errors that are unrelated to the code itself (in other words, your code might be flawless, but the document still will not knit). This can mean additional troubleshooting, and may be a reason to move your code to an R script while you work.

For more in-depth information on R Markdown see [the RMarkdown pages from RStudio](https://rmarkdown.rstudio.com/articles_intro.html).

---

### Writing your thesis in `thesisdown`

The `thesisdown` package provides an R Markdown template for writing your senior thesis. This is particularly helpful for easily integrating R code and graphics into your thesis manuscript. Instructions for how to install this package and use the template can be found in the [GitHub repository](https://github.com/ismayc/thesisdown).

---

### Meet the Palmer Penguins

These tutorials make use of a dataset named `penguins` from the `palmerpenguins` package. The dataset contains different body measurements for three species of penguins from three islands in the Palmer Archipelago, Antarctica.

The `penguins` dataset is useful for learning R, because it contains multiple kinds of data (both categorical and numeric variables). More information about the package and its data can be found on its [github repository](https://github.com/allisonhorst/palmerpenguins) and  [documentation](https://allisonhorst.github.io/palmerpenguins/articles/intro.html). You can read about why we are not using `iris`, another common example dataset, in [this blogpost by Megan Stodel](https://www.meganstodel.com/posts/no-to-iris/). 

```{r, echo = FALSE}
url <- "https://github.com/allisonhorst/palmerpenguins/raw/master/man/figures/lter_penguins.png"

knitr::include_graphics(url)
```
*Artwork by @allison_horst*

If you want to work along with any of the examples on this site, you will first need to install the palmerpenguins package:

```{r, eval = FALSE}
install.packages("palmerpenguins")
```

Note that you only need to _install_ the package once. Once you have installed the package, you will need to _load_ the package every time you want to use the data. To load the package:

```{r}
library(palmerpenguins)
```

Now that the package is loaded, you can access the data with the `data()` function:

```{r}
data("penguins")
```

You might notice that the dataset does not seem to be fully loaded. Next to `penguins`, there is probably a `<Promise>` symbol. To _really_ load the data, click on `<Promise>`. The `penguins` dataset is now ready to use.

You can explore the `penguins` data:

```{r}
penguins
```

In this data, every row is a unique observation (an individual penguin). For each penguin, you have the `species` of the penguin, which `island` it lives on, its `sex`, the `year` the measurements were taken, and some body size measurements.