---
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Meet the Palmer Penguins

Throughout these tutorials, we will be making use of a dataset called `penguins` from the `palmerpenguins` package. The dataset contains various different measurements for penguins of three species on three islands in the Palmer Archipelago, Antarctica.

```{r, echo = FALSE}
url <- "https://github.com/allisonhorst/palmerpenguins/raw/master/man/figures/lter_penguins.png"

knitr::include_graphics(url)
```
*Artwork by @allison_horst*

If you want to work along with any of the examples on this website and need to load the data, you will first need to install the palmerpenguins package:

```{r, eval = FALSE}
install.packages("palmerpenguins")
```

Note that you only need to _install_ the package once. Once you have installed the package, though, you will need to _load_ the package every time you want to use the data. To load the package:

```{r}
library(palmerpenguins)
```

Now you can load the data with the `data()` function:

```{r}
data("penguins")
```

You might notice that the dataset does not seem to be fully loaded in. Beside `penguins`, there is probably a `<Promise>` symbol. To _really_ load the data in, click on `<Promise>`. `penguins` is now ready to go.

We can now explore `penguins`:

```{r}
penguins
```

In this data, every row is a unique observation (penguin). For each penguin, we have the `species` of the penguin, which `island` it lives on, its `sex`, the `year` the measurements were taken, and some body size measurements.

To learn more about all that we can do with this data, check out the remaining tutorials.