---
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Meet the Palmer Penguins

These tutorials make use of a dataset named `penguins` from the `palmerpenguins` package. The dataset contains different body measurements for three species of penguins from three islands in the Palmer Archipelago, Antarctica.

The `penguins` dataset is useful for learning R, because it contains multiple kinds of data (both categorical and numeric variables). More information about the package and its data can be found on its [github repository](https://github.com/allisonhorst/palmerpenguins) and  [documentation](https://allisonhorst.github.io/palmerpenguins/articles/intro.html). You can read about why we are not using `iris`, another common example dataset, in [this blogpost by Megan Stodel](https://www.meganstodel.com/posts/no-to-iris/). 

```{r, echo = FALSE}
url <- "https://github.com/allisonhorst/palmerpenguins/raw/master/man/figures/lter_penguins.png"

knitr::include_graphics(url)
```
*Artwork by @allison_horst*

If you want to work along with any of the examples on this site, you will first need to install the palmerpenguins package:

```{r, eval = FALSE}
install.packages("palmerpenguins")
```

Note that you only need to _install_ the package once. Once you have installed the package, you will need to _load_ the package every time you want to use the data. To load the package:

```{r}
library(palmerpenguins)
```

Now that the package is loaded, you can access the data with the `data()` function:

```{r}
data("penguins")
```

You might notice that the dataset does not seem to be fully loaded. Next to `penguins`, there is probably a `<Promise>` symbol. To _really_ load the data, click on `<Promise>`. The `penguins` dataset is now ready to use.

You can explore the `penguins` data:

```{r}
penguins
```

In this data, every row is a unique observation (an individual penguin). For each penguin, you have the `species` of the penguin, which `island` it lives on, its `sex`, the `year` the measurements were taken, and some body size measurements.
