---
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Meet the Palmer Penguins

These tutorials make use of a dataset named `penguins` from the `palmerpenguins` package. The dataset contains various different measurements for three species of penguins on three islands in the Palmer Archipelago, Antarctica.

The `penguins` dataset is excellent for learning in R, because it contains both categorical and numeric variables, allowing you to practice skills such as calculating statistics and creating data visualizations. More information about the package and its data can be found on its [github repository](https://github.com/allisonhorst/palmerpenguins) or through some additional [documentation](https://allisonhorst.github.io/palmerpenguins/articles/intro.html). You can read about why we are not using `iris`, another common example dataset, in [this blogpost by Megan Stodel](https://www.meganstodel.com/posts/no-to-iris/). 

```{r, echo = FALSE}
url <- "https://github.com/allisonhorst/palmerpenguins/raw/master/man/figures/lter_penguins.png"

knitr::include_graphics(url)
```
*Artwork by @allison_horst*

If you want to work along with any of the examples on this site, you will first need to install the palmerpenguins package:

```{r, eval = FALSE}
install.packages("palmerpenguins")
```

Note that you only need to _install_ the package once. Once you have installed the package, you will need to _load_ the package every time you want to use the data. To load the package:

```{r}
library(palmerpenguins)
```

Now you can load the data with the `data()` function:

```{r}
data("penguins")
```

You might notice that the dataset does not seem to be fully loaded in. Next to `penguins`, there is probably a `<Promise>` symbol. To _really_ load the data in, click on `<Promise>`. `penguins` is now ready to go.

You can now explore `penguins`:

```{r}
penguins
```

In this data, every row is a unique observation (penguin). For each penguin, you have the `species` of the penguin, which `island` it lives on, its `sex`, the `year` the measurements were taken, and some body size measurements.
