---
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Troubleshooting


#### First steps


#### Getting information about a function in R

You can always look up a function name on the internet and read it's documentation, but most R-packages and functions have help files that can be accessed through the RStudio Console.

For example if you can't quite remember what arguments should be used for a function you can type: `?function_name`, into the console and hit enter. This will load up the help file for that function in the RStudio Help panel which has several useful pieces of information:

  - Usage: a general description of the function as well as it's default arguments
  - Arguments: a more detailed breakdown of the function's inputs/arguments. This is particularly useful for finding out       what types of objects the function accepts as arguments.
  - Value: A description of what the output of the function will be
  - Examples: Some code that shows off a common way to use the function.
  
For example, running

```{r}
?mean
```

will bring up the help file. A fairly common problem with using the `mean` function is to run something like:

```{r}
mean(1,2,3)
```

which will not break, but it will output 1 which we know is not the correct value. By looking at the help file we can figure out that the values that we want to find the mean of should all be in the first argument. We can see this both in the "Usage" section and also the "Examples' section. So what we should actually be doing is:

```{r}
mean(c(1,2,3))
```

which works properly.

#### Troubleshooting variable types

A fairly common error message to get is something along the lines of

"argument is not ____ " or " non- ____ argument to a ___ function" or "NAs introduced by coercion"

These often have to do with an argument or variable being the wrong type. For example if you might be trying to perform a mathematical operation of a character vector. There are a couple of good ways to check variables types:

Using `class` we can check the type of vectors and values. For example the output of running the following chunk will be "numeric", "character", and "logical"

```{r}
class(1)
class("a")
class(NA)
```

This can be a bit tedious if you're working with a dataframe though, in which case it's easier to run something like

```{r}
str(data_frame_name)
```

which will give a breakdown of all the variables and their corresponding types.

Now in order to change a variable type, the suite of as.* functions is quite useful. Generally the syntax is `as.desired_type(object_that_needs_converting)`

For example if we get stuck with a "1" instead of a 1, we could use 

```{r}
character_one <- "1"

numeric_one <- as.numeric(character_one)
```

Inside of a dataframe this often looks something like this:

```{r}
data_frame_converted <- old_data_frame %>% 
  mutate(converted_type = as.numeric(old_object))
```

This ties in nicely to the previous section where we detailed how to access the help files of a function. Help files will almost always describe what variable types are expected and so if you're trying to troubleshoot a function that isn't working, checking the help file and your argument types is always a good place to start.

#### Locating the problem in a chunk of code


#### Being aware of NAs







