---
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(ggplot2)
library(knitr)
library(kableExtra)
library(waffle)
```

#### Barplots

Barplots demonstrate the distribution of data across categories (categorical variables) and use the `geom_bar()` function. 

```{r}
ggplot(data = penguins, mapping = aes(x = island)) +
  geom_bar()
```

You can display another variable by including `fill = ` inside of your `aes()` call.

```{r}
ggplot(data = penguins, mapping = aes(x = island, fill = sex)) +
  geom_bar()
```

In the above examples, `ggplot2` determines the height of bars by counting the number of observations behind the scenes. This count provides `geom_bar()` with a `y` aesthetic, so it is not necessary to specify one in the `ggplot2` code.

Your data may already contain counts, in which case you would use a slightly different `geom_` to contruct your barplot. Continuing to work with the penguins example, first create a new dataset with the count of penguins by `island`:

```{r, message = FALSE, warning = FALSE}
penguin_sum <- penguins %>%
  count(island)

penguin_sum
```

To make the same graph as above with pre-counted data, use the `geom_col()` (as in "column") function, rather than `geom_bar()`, and supply the height of the bars explicitly.

```{r}
ggplot(data = penguin_sum, mapping = aes(x = island, y = n)) +
  geom_col()
```

Notice that we had to supply `y = n` inside of `aes()`. A helpful way to remember the difference between `geom_bar()` and `geom_col()` is that Col needs Counts. For additional examples of `geom_bar()` and `geom_col()`, see the barplots section of [ModernDive](https://moderndive.com/2-viz.html#barplots-via-geom_bar-or-geom_col).