---
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(ggplot2)
library(knitr)
library(kableExtra)
library(waffle)
```

```{r, echo = FALSE, message=FALSE, warning=FALSE}
penguins_sum <- penguins %>%
  filter(!is.na(body_mass_g)) %>%
  group_by(island, year) %>%
  summarize(mean_body_mass_g = mean(body_mass_g))
```

### Making informative and readable graphs

In addition to `x`, `y`, `color`, and `fill` covered in the [Intro to data visualization](link) section, you can use additional visual variables to represent variables in your data. Below are some additional methods for conveying more information on your graphs, while still keeping your graphs readable

#### Aesthetic mappings

The aesthetics `shape` and `linetype` can be used to display a categorical variable, while `size` can show a continuous variable. These are used in the same way as `color` above. For example, you could make the same exact graph that was shown in the [Linegraphs](link) section, with different islands shown by `linetype` instead of `color`:

```{r, message = FALSE, warning = FALSE}
ggplot(data = penguins_sum, 
       mapping = aes(x = year, y = mean_body_mass_g, linetype = island)) +
  geom_line()
```

To see more specifics of mapping aesthetics, [this blog post](https://www.datanovia.com/en/blog/ggplot-aes-how-to-assign-aesthetics-in-ggplot2/#group-and-line-type) provides helpful examples and highlights cases where different aesthetic mappings might be particularly appropriate.

There are many good resources available for using `ggplot2` in visualization, including the comprehensive [ggplot2 documentation](https://ggplot2.tidyverse.org/) and the related chapter in [R for Data Science](https://r4ds.had.co.nz/data-visualisation.html).

#### Adjusting transparency with alpha

Sometimes, when using `geom_point()`, points can get a bit crowded. (This is called overplotting.) One way to address this is to change the transparency of the points. In this code, `alpha = 0.5` specifies the transparency of the points as 0.5, on a scale from 0 (most transparent) to 1 (least transparent):

```{r, warning = FALSE}
ggplot(data = penguins, mapping = aes(x = bill_length_mm, y = bill_depth_mm, color = sex)) +
  geom_point(alpha = 0.5)
```

#### Creating side-by-side graphs (faceting)

To add another variable to an already "busy" graph, `facet_wrap()` can be useful. Consider the following plot:

```{r, warning = FALSE}
ggplot(data = penguins, mapping = aes(x = bill_length_mm, y = bill_depth_mm, color = sex)) +
  geom_point(alpha = 0.7)
```

Incorporating `facet_wrap()` separates the data by species and plots them side-by-side, providing a visual that shows how bill length and bill depth vary across species _and_ across sex:

```{r, warning = FALSE}
ggplot(data = penguins, mapping = aes(x = bill_length_mm, y = bill_depth_mm, color = sex)) +
  geom_point(alpha = 0.7) +
  facet_wrap(~species)
```

Faceting by `species` creates three graphs, with identical x and y axis scales and coloring. These side-by-side graphs facilitate comparisons across values of `species`.