---
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(ggplot2)
library(knitr)
library(kableExtra)
library(waffle)
```

#### Linegraphs

Linegraphs also show the relationship between two continuous variables, often showing progression over time. You can create a linegraph in much the same way as a scatterplot by swapping out `geom_point()` for `geom_line()`.

The below code is very similar to that in the [Scatterplots](link) section, with two changes: (1) time (year) is now on the x-axis and (2) a different `geom` is being used to represent the same data. 

```{r, message = FALSE, warning = FALSE}
ggplot(data = penguins, 
       mapping = aes(x = year, y = body_mass_g, color = island)) +
  geom_line()
```

The above graph looks a bit odd; this is because there are multiple data points (body mass, y) at each time point (year, x). Linegraphs work best when your data has one value per x location. In this example, that means you would want one y value for _each_ island in _each_ year. You can achieve this by taking the mean of `body_mass_g` for each combination of `island` and `year`:  
```{r, message=FALSE, warning=FALSE}
penguins_sum <- penguins %>%
  filter(!is.na(body_mass_g)) %>%
  group_by(island, year) %>%
  summarize(mean_body_mass_g = mean(body_mass_g))

penguins_sum
```
(In the above code, note the removal of missing values for `body_mass_g` before calculating the mean body mass.)

Now you can make the same linegraph, using the new `penguins_sum` dataset, and with `mean_body_mass_g` in place of `body_mass_g`

```{r, message = FALSE, warning = FALSE}
ggplot(data = penguins_sum, 
       mapping = aes(x = year, y = mean_body_mass_g, color = island)) +
  geom_line()
```


You may come across problems with your linegraphs when your x-variable is categorical. This issue can be resolved by adding `group = 1` into your `aes()` call.


