---
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(knitr)
library(kableExtra)
library(waffle)
```

# Presenting and Visualizing Data

### Intro to data visualization with `ggplot2`

`ggplot2` is an R package for data visualization that makes appealing and effective graphs using a standardized syntax.

To run these examples on your own, load `palmerpenguins`, a package that contains a dataset about penguin populations on islands in Antarctica. You also need to load the `ggplot2` package, which is contained _within_ the `tidyverse` package. (You used other `tidyverse` libraries in the [Transforming data with `dplyr`](link) section.) You can also load `ggplot2` on its own, using `library(ggplot2)`.

To learn more about packages and the `palmerpenguins` package, see our [Getting Started](LINK-HERE) page.

```{r, eval = FALSE}
library(palmerpenguins)
library(tidyverse)
```

Before graphing using `ggplot2`, the data needs to be "tidy". (Read more about tidy data on [our data wrangling page](LINK-HERE) or [from the tidyverse team](https://cran.r-project.org/web/packages/tidyr/vignettes/tidy-data.html). For this example, the `palmerpenguins` data is already tidy.

There are three main pieces that make up a ggplot:

1. Data: You have to specify the data that you want `ggplot()` to plot using the following syntax: 
      
              data = ___

2. Aesthetics: In order for `ggplot()` to know which variables you want to use you have to specify what they are by using some combination of aesthetics, such as:

              aes(x = ___ , y = ___ , color = ___ , fill = ___ )

Anything that you put inside of `aes()` should be a variable. The aesthetics that you use will vary with graph type. (For example, a histogram shows the distribution of one variable and will include `x = ___ `, and not all of the aesthetics listed above; a bar plot may use `fill = `, but a scatterplot would use `color = `.)

3. Geometry: You can think of the `geom` as specifying what shape your data will take, appearing in `ggplot2` code as variations on:

              geom_ ___()

The `ggplot2` package includes a number of standard geometries. Some common `geoms` include `geom_point()` for a scatterplot, `geom_histogram()` for a histogram, and `geom_boxplot()` for a boxplot. 

You can use these three pieces to create a template for making graphs with `ggplot()`:

              ggplot(data = ___ ,
                     mapping = aes(x = ___  , y = ___  , color = ___  , fill = ___   ) +
                geom_ ___()
          
Note that in `ggplot2`, layers are added to the graph (using `+`) rather than the pipe `%>%`. When debugging `ggplot2` code, make sure you have added layers rather than piped together components.

The next several sections provide examples of these commonly-used data visualizations:

  - [Scatterplots](INTERNAL-LINK)
  - [Linegraphs](INTERNAL-LINK)
  - [Barplots](INTERNAL-LINK)
  - [Histograms](INTERNAL-LINK)
  - [Boxplots](INTERNAL-LINK)
  
---

#### Scatterplots

Scatterplots are commonly used to visualize the relationship between two continuous variables. For example, you may be interested in the relationship between body mass and bill length _across_ all islands:

```{r, message=FALSE, warning=FALSE}
ggplot(data = penguins, 
       mapping = aes(x = bill_length_mm, y = body_mass_g)) +
  geom_point()
```

The above code first specifies the `data` (`penguins`). The `mapping` arguments define "which variables belong where". In this case, `bill_length_mm` is on the x axis, `body_mass_g` on the y. The `geom` of `geom_point` specifies that this is a scatterplot.

You could add the visual variable of color to show patterns _within_ each island:

```{r, message=FALSE, warning=FALSE}
ggplot(data = penguins, 
       mapping = aes(x = bill_length_mm, y = body_mass_g, color = island)) +
  geom_point()
```

---

#### Linegraphs

Linegraphs also show the relationship between two continuous variables, often showing progression over time. You can create a linegraph in much the same way as a scatterplot by swapping out `geom_point()` for `geom_line()`.

The below code is very similar to that in the [Scatterplots](link) section, with two changes: (1) time (year) is now on the x-axis and (2) a different `geom` is being used to represent the same data. 

```{r, message = FALSE, warning = FALSE}
ggplot(data = penguins, 
       mapping = aes(x = year, y = body_mass_g, color = island)) +
  geom_line()
```

The above graph looks a bit odd; this is because there are multiple data points (body mass, y) at each time point (year, x). Linegraphs work best when your data has one y value per x value In this example, that means you would want one value of `body_mass_g` for _each_ island in _each_ year. You can achieve this by taking the mean of `body_mass_g` for each combination of `island` and `year`:  

```{r, message=FALSE, warning=FALSE}
penguins_sum <- penguins %>%
  filter(!is.na(body_mass_g)) %>%
  group_by(island, year) %>%
  summarize(mean_body_mass_g = mean(body_mass_g))

penguins_sum
```

(In the above code, note the removal of missing values for `body_mass_g` before calculating the mean body mass.)

Now you can make the same linegraph, using the new `penguins_sum` dataset, and with `mean_body_mass_g` in place of `body_mass_g`

```{r, message = FALSE, warning = FALSE}
ggplot(data = penguins_sum, 
       mapping = aes(x = year, y = mean_body_mass_g, color = island)) +
  geom_line()
```

You may encounter problems with your linegraphs when your x-variable is categorical. This issue can be resolved by adding `group = 1` into your `aes()` call.

---

#### Barplots

Barplots demonstrate the distribution of data across categories (categorical variables) and use either the `geom_bar()` or `geom_col()` function. 

```{r}
ggplot(data = penguins, mapping = aes(x = island)) +
  geom_bar()
```

You can display another variable by including `fill = ` inside of your `aes()` call.

```{r}
ggplot(data = penguins, mapping = aes(x = island, fill = sex)) +
  geom_bar()
```

In the above examples, `ggplot2` determines the height of bars by counting the number of observations (this counting is done automatically and is not specified in the `ggplot` call). This count provides `geom_bar()` with a `y` aesthetic, so it is not necessary to specify one in the `ggplot2` code.

Your data may already contain counts, in which case you would use a different `geom_` to contruct your barplot (`geom_col`). Continuing to work with the penguins example, first create a new dataset with the count of penguins by `island`:

```{r, message = FALSE, warning = FALSE}
penguin_sum <- penguins %>%
  count(island)

penguin_sum
```

To make the same graph as above with pre-counted data, use the `geom_col()` (as in "column") function, and specify the height of the bars.

```{r}
ggplot(data = penguin_sum, mapping = aes(x = island, y = n)) +
  geom_col()
```

Notice that you had to specify `y = n` inside of `aes()`. A helpful way to remember the difference between `geom_bar()` and `geom_col()` is that Col needs Counts. For additional examples of `geom_bar()` and `geom_col()`, see the barplots section of [ModernDive](https://moderndive.com/2-viz.html#barplots-via-geom_bar-or-geom_col).

---

#### Histograms

Histograms show the distribution of data for a continuous variable, specified with `geom_histogram()`.

```{r, message=FALSE, warning=FALSE}
ggplot(data = penguins, mapping = aes(x = body_mass_g)) +
  geom_histogram()
```

Changing either the number of bins, or the width of those bins, can greatly affect the resulting visual. Compare the two below graphs to the initial histogram (notice you can specify `bins` _or_ `binwidth`).

```{r, warning = FALSE}
ggplot(data = penguins, mapping = aes(x = body_mass_g)) +
  geom_histogram(bins = 100)
```
```{r, warning = FALSE}
ggplot(data = penguins, mapping = aes(x = body_mass_g)) +
  geom_histogram(binwidth = 200)
```

When creating a histogram of your data, you will likely want to try different binning schemes to see what provides the best representation of your data. Generally you want to avoid histograms that look like barcodes as well as ones that look like a single block.
---

#### Boxplots

Boxplots can be another useful tool for visualizing a dataset, showing both the spread of the data (range of most of the data, presence of outliers) and how the data is concentrated in the middle (mean, median). 

Boxplots are often most useful to compare data across categories. For example, you can construct a boxplot for penguin weight (body mass in g) across _all_ islands:

```{r, message=FALSE, warning=FALSE}
ggplot(data = penguins, mapping  = aes(y = body_mass_g)) +
  geom_boxplot()
```

But it may be more informative to break that data apart by island: 

```{r, message=FALSE, warning=FALSE}
ggplot(data = penguins, mapping  = aes(x = island, y = body_mass_g)) +
  geom_boxplot()
```

You can create multiple boxplots on the same graph by setting `x` to a variable with multiple categories. As seen in the previous examples, you can also use `color` to illustrate additional variables within one visual:

```{r, message=FALSE, warning=FALSE}
ggplot(data = penguins, mapping  = aes(x = island, y = body_mass_g, color = sex)) +
  geom_boxplot()
```

(The dots you see as part of the boxplots are outliers, which `geom_boxplot` defines as being outside of 150% the value of the interquartile range [IQR], the difference between the 1st [25%] and the 3rd [75%] quartile of your data.)

A common problem when creating multiple boxplots on the same graph is that even though you supply an x-variable, your plot only has one boxplot. This usually happens when the x-variable is not being properly treated as a factor. One option for a fix is to encode the variable as a factor within the `ggplot` call:

```{r}
ggplot(data = penguins, mapping  = aes(x = factor(island), y = body_mass_g)) +
  geom_boxplot()
```

---

### Making informative and readable graphs

In addition to `x`, `y`, `color`, and `fill` covered in the [Intro to data visualization](link) section, you can use additional visual variables to represent variables in your data. Below are some additional methods for conveying more information on your graphs, while still keeping your graphs readable

#### Aesthetic mappings

The aesthetics `shape` and `linetype` are best used to display a categorical variable, while `size` is best used with continuous variables. These are used in the same way as `color` above. For example, you could make the same graph that was shown in the [Linegraphs](link) section, with different islands shown by `linetype` instead of `color`:

```{r, message = FALSE, warning = FALSE}
ggplot(data = penguins_sum, 
       mapping = aes(x = year, y = mean_body_mass_g, linetype = island)) +
  geom_line()
```

To see more specifics of mapping aesthetics, [this blog post](https://www.datanovia.com/en/blog/ggplot-aes-how-to-assign-aesthetics-in-ggplot2/#group-and-line-type) provides helpful examples and highlights cases where different aesthetic mappings might be particularly appropriate.

There are many resources available for using `ggplot2` in visualization, including the comprehensive [ggplot2 documentation](https://ggplot2.tidyverse.org/) and the related chapter in [R for Data Science](https://r4ds.had.co.nz/data-visualisation.html).

#### Adjusting transparency with alpha

Sometimes, when using `geom_point()`, points can get a bit crowded. (This is called overplotting.) One way to address this is to change the transparency of the points. In the code below, `alpha = 0.5` specifies the transparency of the points as 50%, on a scale from 0 (most transparent) to 1 (least transparent):

```{r, warning = FALSE}
ggplot(data = penguins, mapping = aes(x = bill_length_mm, y = bill_depth_mm, color = sex)) +
  geom_point(alpha = 0.5)
```

#### Creating side-by-side graphs (faceting)

To add another variable to an already "busy" graph, `facet_wrap()` can be useful. Consider the following plot:

```{r, warning = FALSE}
ggplot(data = penguins, mapping = aes(x = bill_length_mm, y = bill_depth_mm, color = sex)) +
  geom_point(alpha = 0.7)
```

Incorporating `facet_wrap()` separates the data by species and plots them side-by-side, providing a visual that shows how bill length and bill depth vary across species _and_ across sex:

```{r, warning = FALSE}
ggplot(data = penguins, mapping = aes(x = bill_length_mm, y = bill_depth_mm, color = sex)) +
  geom_point(alpha = 0.7) +
  facet_wrap(~species)
```

Faceting by `species` creates three graphs, with identical x and y axis scales and coloring. These side-by-side graphs facilitate comparisons across values of `species`.

---

### Reordering Categories in a Plot

When building plots with categorical data, `ggplot` will default to ordering categories alphabetically; this may not make sense for your data. (Example: a barplot showing days of the week on the x-axis; alphabetical order is not the correct ordering/sequencing.) Within the `forcats` package, there is a family of functions that helps with these types of problems.

#### Reordering by another variable

##### Barplots

Make a plot comparing how many penguins of each species are in the `palmerpenguins` dataset. Notice that the bars are ordered alphabetically, when perhaps you would rather have them ordered by `n`, the number of penguins of each species.

```{r}
penguins %>% 
  count(species) %>% 
  ggplot(aes(x = species, y = n)) +
  geom_col()
```

If you wish to order a categorical variable (`species`) by some quantitative variable (`n`) `fct_reorder()` is a useful option. Generally the syntax is as follows:

```
fct_reorder(.f = categorical_variable_to_be_ordered, .x = quantitative_variable_to_order_other_variable_by)
```

To make a change to the "factor levels" of the data, use this function inside of `mutate`:

```{r}
penguins %>% 
  count(species) %>% 
  mutate(species_ordered = fct_reorder(.f = species, .x = n)) %>%
  ggplot(aes(x = species_ordered, y = n)) +
  geom_col()
```

Note the change in the order of the bars in the graph produced by the code below. To order the bars in descending rather than ascending (numeric) order, set the `.desc` argument to TRUE in `fct_reorder()`:

```{r}
penguins %>% 
  count(species) %>% 
  mutate(species_ordered = fct_reorder(.f = species, .x = n, .desc = TRUE)) %>%
  ggplot(aes(x = species_ordered, y = n)) +
  geom_col()
```

##### Boxplots

Boxplots can be trickier because there may not be some explicit variable in the dataset that `fct_reorder` can use as a reference. You can use `fct_reorder` to order boxplots by some function, which is generally a summary function that can be applied to your y-variable. (A few examples include `max`, `min`, `median`, and `mean`.)

For example, start with a plot featuring multiple boxplots (one for each species), where the boxplots are ordered alphabetically:

```{r}
penguins %>% 
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_boxplot()
```

For contrast, create a plot with multiple boxplots ordered by the median value of `bill_length_mm`.

The below code uses the argument `.fun` to reorder by median. (Note that the function is written as `median` and not as `median()`.) You may notice the specification `na.rm = T`; this removes missing data from the calculation of median. If there are missing values in your data, it may cause the reordering to not function properly. 

```{r}
penguins %>% 
  mutate(species = fct_reorder(.f = species, .x = bill_length_mm, .fun = median, na.rm = T)) %>% 
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_boxplot()
```

Similar to the earlier example, you can put groups in descending order using `.desc`:

```{r}
penguins %>% 
  mutate(species = fct_reorder(.f = species, .x = bill_length_mm, .fun = median, na.rm = T, .desc = T)) %>% 
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_boxplot()
```

#### Reordering manually

##### Barplots

For a different example, create a new demo dataset called `days_of_week`, showing the hours worked on each day:

```{r}
days_of_week <- tibble(
  days = c('Fri', 'Sat', 'Sun', 'Mon', 'Tue', 'Wed', 'Thu'),
  hours_worked = seq(20,50, 5)
)

days_of_week
```

Plotting this data using a barplot yields:

```{r}
days_of_week %>% 
  ggplot(aes(x = days, y = hours_worked)) +
  geom_col()
```

The columns are automatically sorted alphabetically, but days of the week are _ordinal data_, meaning that the order of the values is important. Here `fct_reorder` is not the correct tool because order is not defined by another variable, but by external information. (In this case, knowledge of a calendar.)

The `forcats` function `fct_relevel` will help here. The syntax is generally:

```
fct_relevel(.f = factor_that_needs_reordering, levels = levels_to_order_by)
```

The below code creates a vector with days (levels) in the desired order, and then feeds that vector into the `fct_relevel` call.

```{r}
day_order <- c('Sun', 'Mon', 'Tue', 'Wed', 'Thu', 'Fri', 'Sat')

days_of_week %>% 
  mutate(days = fct_relevel(.f = days, levels = day_order)) %>% 
  ggplot(aes(x = days, y = hours_worked)) +
  geom_col()

```

After reordering factors, the bar order in the bargraph has changed. 

##### Boxplots

To see how this works with Boxplots, make some changes to the demo dataset:

```{r}
days_of_week_new <- tibble(
  days = rep(c('Fri', 'Sat', 'Sun', 'Mon', 'Tue', 'Wed', 'Thu'), 10),
  hours_worked = rnorm(70, 25, 7)
)
```

Plotting once again yields a x-axis ordered alphabetically, which is not what we want. 

```{r}
days_of_week_new %>% 
  ggplot(aes(x = days, y = hours_worked)) +
  geom_boxplot()
```

Use the same vector describing the levels inside of `fct_relevel` to get:

```{r}
days_of_week_new %>% 
  mutate(days = fct_relevel(.f = days, levels = day_order )) %>% 
  ggplot(aes(x = days, y = hours_worked)) +
  geom_boxplot()
```

---

### Extensions of `ggplot2`

#### More graph types and add-ons

Although some of the most frequently used graph types (geoms) are described above, some other commonly used geoms include:

  - `geom_density()`: for [density estimates](https://ggplot2.tidyverse.org/reference/geom_density.html) 
  - `geom_dotplot()`: for [dot-plots](https://ggplot2.tidyverse.org/reference/geom_dotplot.html?q=dotplot)
  - `geom_violin()`: for making [violin plots](https://ggplot2.tidyverse.org/reference/geom_violin.html)
  - `geom_jitter()`: for dealing with [overplotting](https://ggplot2.tidyverse.org/reference/geom_jitter.html).

There are also many extra add-ons (layers) that you can add to your plots using the following functions:

  - `geom_abline()`, `geom_hline()`, `geom_vline()`: for adding [customizable reference lines](https://ggplot2.tidyverse.org/reference/geom_abline.html?q=geom%20_%20hli#arguments)
  - `geom_errorbar()`: for adding [error bars](https://ggplot2.tidyverse.org/reference/geom_linerange.html)
  - `geom_smooth()`: for adding [regression lines](https://ggplot2.tidyverse.org/reference/geom_smooth.html?q=smooth) and other conditional means
  
#### Waffle plots with `waffle`

Waffle plots (sometimes referred to as "square pie charts") are a way of visualizing category breakdowns, where a grid of boxes is filled in with different colors to match your categories. Waffle plots can be created in R with the [`waffle` package](https://cran.r-project.org/web/packages/waffle/waffle.pdf). You can find examples of how to use the `waffle()` function to make these plots [on the R Documentation site](https://www.rdocumentation.org/packages/waffle/versions/0.7.0/topics/waffle) or on [R Bloggers](https://www.r-bloggers.com/2015/03/making-waffle-charts-in-r-with-the-new-waffle-package/).

#### Mosaic plots with `ggmosaic`

Mosaic plots are a way of visually displaying data from two or more categorical variables. The area of cells are used to show the frequency of observations. Mosaic plots can be created in R with the [`ggmosaic` package](https://cran.r-project.org/web/packages/ggmosaic/vignettes/ggmosaic.html), or the [vcd package](https://r-graphics.org/recipe-miscgraph-mosaic). 

#### Spatial data with `ggmap`

You can use R for a wealth of spatial analyses (not covered here), and can also use R to make quick maps. `ggmap` is a package designed to follow many of the same basic principles and syntax as `ggplot2`, that allows you to display spatial data. The [`ggmap` CRAN site](https://cran.r-project.org/web/packages/ggmap/readme/README.html) has some helpful example maps, as does [the R Graph Gallery](https://www.r-graph-gallery.com/324-map-background-with-the-ggmap-library.html).

---

### Graph customization 

Once you have a basic graph, you may want to further alter its appearance. Some common customizations include:

- changing axis labels, titles, and captions (using `labs()`)
- specifying colors within the graph (e.g. with `scale_fill_manual`)
- applying a `theme()` to the graph, changing its overall appearance (e.g. `theme_dark()`, `theme_bw()`, `theme_classic()`,`theme_minimal()`
  
You can also change individual aspects of your them inside of the `theme()` call. For specifics of working with `theme()`, see [the ggplot2 documentation](https://ggplot2.tidyverse.org/reference/theme.html).

To see these customizations in action, start with the scatterplot showing penguin bill length and body mass by island (created above), and add a few lines to change title, color, and labels:

```{r, message=FALSE, warning=FALSE}
ggplot(data = penguins, 
       mapping = aes(x = bill_length_mm, y = body_mass_g, color = island)) +
  geom_point(alpha = 0.7) +
  scale_color_manual(values = c("magenta3", "darkorange", "blue4")) +
  labs(x = "Bill length (mm)",
       y = "Body mass (g)",
       title = "Palmer Penguins",
       subtitle = "Examining penguin body mass and bill length") + 
  theme_bw()
  
```

To read more about `ggplot2`, see the [R for Data Science](https://r4ds.had.co.nz/data-visualisation.html) chapter about data visualization. More extensive documentation can be found on the [ggplot2 website](https://ggplot2.tidyverse.org/).
---

### Summary tables

#### kable and kableExtra

The [Summarizing data](https://www.reed.edu/data-at-reed/resources/R/summarizing-data.html) page shows how to compute summaries of your data. Summary tables can be useful for displaying data, and the `kable()` function in the R package `knitr` allows you to present tables with helpful formatting.

First, install and load the package.

```{r eval = FALSE}
install.packages("knitr")
library(knitr)
```

Using the `palmerpenguins` data, calculate the mean value of penguin body mass (g) across islands over time. (For more information on calculating summary statistics, see the [Data Wrangling](https://www.reed.edu/data-at-reed/resources/R/transforming-data-with-dplyr.html)) section.)  

The below code calculates summary statistics, saved in a new dataset called `penguin_sum`.  

```{r message = FALSE}
penguin_sum <- penguins %>%
  group_by(island, year) %>%
  summarize(mean_body_mass_g = mean(body_mass_g, na.rm = TRUE)) %>% 
  ungroup()
```

You can print the dataset within R/RStudio:

```{r}
penguin_sum
```

To make a possibly more appealing table with the same data, use `kable()`:

```{r}
kable(x = penguin_sum, 
      format = "html",
      col.names = c("Island", "Year", "Mean Body Mass (g)"),
      caption = "Mean body mass of penguins on different islands over time")
```

For additional customization options, use the package `kableExtra`. 

```{r eval = FALSE}
install.packages("kableExtra")
library(kableExtra)
```

Start with very similar code to what you used with the basic `kable` table. Add the function `kable_styling()` to get the basic `kableExtra` format:

```{r}
kable(x = penguin_sum, 
      col.names = c("Island", "Year", "Mean Body Mass (g)"),
      caption = "Mean body mass of penguins on different islands over time") %>%
  kable_styling()
```

Some other options to consider as alternatives to `kable_styling()` are: `kable_classic()`, `kable_paper()`, `kable_classic_2()`,`kable_minimal()`, `kable_material()` and `kable_material_dark()`. 

To change the width of the table, add `full_width = FALSE` within your `kable_styling()` (or similar) argument. Additional options with `kableExtra` include changing the font:

```{r}
kable(x = penguin_sum, 
      col.names = c("Island", "Year", "Mean Body Mass (g)"),
      caption = "Mean body mass of penguins on different islands over time") %>%
   kable_classic(full_width = FALSE, html_font = "Cambria", font_size=16)
```

You can also adjust the font size: 
```{r}
kable(x = penguin_sum, 
      col.names = c("Island", "Year", "Mean Body Mass (g)"),
      caption = "Mean body mass of penguins on different islands over time") %>%
   kable_material_dark(html_font = "Cambria")
```

To learn more about these and other customizations, see [the kableExtra vignette](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#Overview).


#### gt

Another approach for creating tables is the `gt` package. `gt` stands for "grammar of tables," and feels a lot like working with `ggplot`'s grammar of graphics.

```{r}
install.packages("gt")
library(gt)
```

You can create a basic table by adding `gt()` on to the end of your code:

```{r}
penguin_sum %>% 
  gt()
```

You can also adjust your column names and include a title by adding on to your baseline table. Notice that you can use the `md()` wrapper, which stands for "markdown," to use markdown styling in your tables; the `**` on either side will bold the `title` argument. One `*` on either side will italicize the text.

```{r}
penguin_sum %>% 
  gt() %>% 
  cols_label(
    mean_body_mass_g = "Mean Body Mass (g)"
    ) %>% 
  tab_header(
    title = md("**Mean body mass of penguins on different islands**"),
    subtitle = "2007-2009"
    ) 
```

Finally, you can both group and color your columns to help the viewer differentiate between the species. Before adding color to your table, you will need to complete a few additional steps.

First, install a few new packages and choose a color palette from the many available: 

```{r}
install.packages("paletteer")
install.packages("scales")
```

You can run this line of code to look at all the color options:

```{r, eval = FALSE}
info_paletteer(color_pkgs = NULL)
```

The syntax is generally `packagename::palette` and you will usually have to install the package before accessing the palette. Notice that you can specify how many colors we want with the `n = 3` argument.

```{r}
install.packages("ggsci")

palette <- paletteer::paletteer_d("ggsci::teal_material", n = 3) %>%
  as.character()
```

```{r}
penguin_sum %>% 
  gt() %>% 
  cols_label(
    mean_body_mass_g = "Mean Body Mass (g)"
    ) %>% 
  tab_header(
    title = md("**Mean body mass of penguins on different islands**"),
    subtitle = "2007-2009"
    ) %>% 
  data_color(
    columns = "island",
    colors = scales::col_factor(
      as.character(palette),
      domain = NULL
    )
  )
```

As with kable, there is a lot that you can do to customize your `gt` tables. These examples do not provide a full introduction to the options available within `gt`; there are many great resources which provide more deatil. For more on using `gt`, consider starting with that go in depth on using the package---[the `gt` website](https://gt.rstudio.com/reference/index.html).
