---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(knitr)
```

### Reordering Categories in a Plot

When building plots with categorical data, `ggplot` will default to ordering categories alphabetically; this may not make sense for your data. (Example: a barplot showing days of the week on the x-axis; alphabetical order is not the correct ordering/sequencing.) Within the `forcats` package, there is a family of functions that helps with these types of problems.

#### Reordering by another variable

##### Barplots

Make a plot comparing how many penguins of each species are in the `palmerpenguins` dataset. Notice that the bars are ordered alphabetically, when perhaps you would rather have them ordered by `n`, the number of penguins of each species.

```{r}
penguins %>% 
  count(species) %>% 
  ggplot(aes(x = species, y = n)) +
  geom_col()
```

If you wish to order a categorical variable (`species`) by some quantitative variable (`n`) `fct_reorder()` is a useful option. Generally the syntax is as follows:

```
fct_reorder(.f = categorical_variable_to_be_ordered, .x = quantitative_variable_to_order_other_variable_by)
```

To make a change to the "factor levels" of the data, use this function inside of `mutate`:

```{r}
penguins %>% 
  count(species) %>% 
  mutate(species_ordered = fct_reorder(.f = species, .x = n)) %>%
  ggplot(aes(x = species_ordered, y = n)) +
  geom_col()
```

Note the change in the order of the bars in the graph produced by the code below. To order the bars in descending rather than ascending (numeric) order, set the `.desc` argument to TRUE in `fct_reorder()`:

```{r}
penguins %>% 
  count(species) %>% 
  mutate(species_ordered = fct_reorder(.f = species, .x = n, .desc = TRUE)) %>%
  ggplot(aes(x = species_ordered, y = n)) +
  geom_col()
```

##### Boxplots

Boxplots can be trickier because there may not be some explicit variable in the dataset that `fct_reorder` can use as a reference. You can use `fct_reorder` to order boxplots by some function, which is generally a summary function that can be applied to your y-variable. (A few examples include `max`, `min`, `median`, and `mean`.)

For example, start with a plot featuring multiple boxplots (one for each species), where the boxplots areordered alphabetically:

```{r}
penguins %>% 
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_boxplot()
```

For contrast, create a plot with multiple boxplots ordered by the median value of `bill_length_mm`.

The below code uses the argument `.fun` to reorder by median. (Note that the function is written as `median` and not as `median()`.) You may notice the specification `na.rm = T`; this removes missing data from the calculation of median. If there are missing values in your data, it may cause the reordering to not function properly. 

```{r}
penguins %>% 
  mutate(species = fct_reorder(.f = species, .x = bill_length_mm, .fun = median, na.rm = T)) %>% 
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_boxplot()
```

Similar to the earlier example, you can put groups in descending order using `.desc`:

```{r}
penguins %>% 
  mutate(species = fct_reorder(.f = species, .x = bill_length_mm, .fun = median, na.rm = T, .desc = T)) %>% 
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_boxplot()
```

#### Reordering manually

For a different example, create a new demo dataset called `days_of_week`, showing the hours worked on each day:

```{r}
days_of_week <- tibble(
  days = c('Fri', 'Sat', 'Sun', 'Mon', 'Tue', 'Wed', 'Thu'),
  hours_worked = seq(20,50, 5)
)

days_of_week
```

Plotting this data using a barplot yields:

```{r}
days_of_week %>% 
  ggplot(aes(x = days, y = hours_worked)) +
  geom_col()
```

The columns are automatically sorted alphabetically, but days of the week are _ordinal data_, meaning that the order of the values is important. Here `fct_reorder` is not the correct tool because order is not defined by another variable, but by external information. (In this case, knowledge of a calendar.)

The `forcats` function `fct_relevel` will help here. The syntax is generally:

```
fct_relevel(.f = factor_that_needs_reordering, levels = levels_to_order_by)
```

The below code creates a vector with days (levels) in the desired order, and then feeds that vector into the `fct_relevel` call.

```{r}
day_order <- c('Sun', 'Mon', 'Tue', 'Wed', 'Thu', 'Fri', 'Sat')

days_of_week %>% 
  mutate(days = fct_relevel(.f = days, levels = day_order)) %>% 
  ggplot(aes(x = days, y = hours_worked)) +
  geom_col()

```

After reordering factors, the bar order in the bargraph has changed. This same approach can be used with boxplots.

Another option is to do the reordering inside of the `ggplot` call:

```{r}
days_of_week %>% 
  ggplot(aes(x = fct_relevel(.f = days, levels = day_order), y = hours_worked)) +
  geom_col()
```

While the above graph shows the data/levels correctly ordered, the underlying levels of the data have not been changed. For consistency across visualizations and analysis, consider using `fct_*` functions inside of `mutate`.









