---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(knitr)
```

### Reordering Categories in a Plot

When making barplots or boxplots with multiple categories it isn't uncommon for ggplot to put your categories in an order that you don't want. Common examples of this are a boxplot that you wish were ordered by height of the bars, or a barplot where the days of the week are displayed out of order. There is a whole family of functions within the `forcats` package that helps with these types of problems.

#### Reordering by another variable

Let's make a plot comparing how many penguins of each species we have in the `palmerpenguins` dataset. Notice that the bars are ordered alphabetically when we'd rather have them ordered by `n`, which is the number of penguins of each species.

```{r}
penguins %>% 
  count(species) %>% 
  ggplot(aes(x = species, y = n)) +
  geom_col()
```

When we want to order a categorical variable (`species`) by some quantitative variable (`n`) `fct_reorder()` is the best option. Generally the syntax is as follows:

```
fct_reorder(categorical_variable_to_be_ordered, quantitative_variable_to_order_other_variable_by)
```

We'll use the function inside of `mutate` since we're changing the underlying levels attached to our data.

```{r}
penguins %>% 
  count(species) %>% 
  mutate(species_ordered = fct_reorder(species, n)) %>%
  ggplot(aes(x = species_ordered, y = n)) +
  geom_col()
```

And as desired, our bars are now in order. If we wanted them to be in descending rather than ascending order we can just add a minus sign in from of the `n` in `fct_reorder()`:

```{r}
penguins %>% 
  count(species) %>% 
  mutate(species_ordered = fct_reorder(species, -n)) %>%
  ggplot(aes(x = species_ordered, y = n)) +
  geom_col()
```


#### Reordering manually

#maybe use days of the week example or "control" "A" "B" treatments here
