---
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(knitr)
library(kableExtra)
library(waffle)
```

#### Scatterplots

Scatterplots are commonly used to visualize the relationship between two continuous variables. For example, you may be interested in the relationship between body mass and bill length _across_ all islands:

```{r, message=FALSE, warning=FALSE}
ggplot(data = penguins, 
       mapping = aes(x = bill_length_mm, y = body_mass_g)) +
  geom_point()
```

The above code first specifies the `data` (`penguins`). The `mapping` arguments define "which variables belong where". In this case, `bill_length_mm` is on the x axis, `body_mass_g` on the y. The `geom` of `geom_point` specifies that this is a scatterplot.

You could add the visual variable of color to show patterns _within_ each island:

```{r, message=FALSE, warning=FALSE}
ggplot(data = penguins, 
       mapping = aes(x = bill_length_mm, y = body_mass_g, color = island)) +
  geom_point()
```