---
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(ggplot2)
library(knitr)
library(kableExtra)
library(waffle)
```

### Summary tables

The [Summarizing data](LINK-HERE) page shows how to compute summaries of your data. Summary tables can be useful for displaying data, and the `kable()` function in the R package `knitr` allows you to present tables with helpful formatting.

First, install and load the package.

```{r eval = FALSE}
install.packages("knitr")
library(knitr)
```

Using the `palmerpenguins` data, calculate the mean value of penguin body mass (g) across islands over time. (For more on calculating summary statistics, see the [Data Wrangling](LINK-HERE) section.)

Here is the code for our summary statistic, which we will save in a new dataset called `penguin_sum`.

```{r message = FALSE}
penguin_sum <- penguins %>%
  group_by(island, year) %>%
  summarise(mean_body_mass_g = mean(body_mass_g, na.rm = TRUE))
```

You can print the dataset within R/RStudio:

```{r}
penguin_sum
```

To make a possibly more appealing table with the same data, use `kable()`:

```{r}
kable(x = penguin_sum, 
      format = "html",
      col.names = c("Island", "Year", "Mean Body Mass (g)"),
      caption = "Mean body mass of penguins on different islands over time")
```



For additional customization options, use the package `kableExtra`. 

```{r eval = FALSE}
install.packages("kableExtra")
library(kableExtra)
```

Start with very similar code to what you used with the basic `kable` table. Add the function `kable_styling()` to get the basic `kableExtra` format:

```{r}
kable(x = penguin_sum, 
      col.names = c("Island", "Year", "Mean Body Mass (g)"),
      caption = "Mean body mass of penguins on different islands over time") %>%
  kable_styling()
```

Some other options to consider as alternatives to `kable_styling()`: `kable_classic()`, `kable_paper()`, `kable_classic_2()`,`kable_minimal()`, `kable_material()` and `kable_material_dark()`. 

To change the width of the table, add `full_width = FALSE` within your `kable_styling()` (or similar) argument. Additional options with `kableExtra` include changing the font:

```{r}
kable(x = penguin_sum, 
      col.names = c("Island", "Year", "Mean Body Mass (g)"),
      caption = "Mean body mass of penguins on different islands over time") %>%
   kable_classic(full_width = FALSE, html_font = "Cambria", font_size=16)
```

You can also adjust the font size: 
```{r}
kable(x = penguin_sum, 
      col.names = c("Island", "Year", "Mean Body Mass (g)"),
      caption = "Mean body mass of penguins on different islands over time") %>%
   kable_material_dark(html_font = "Cambria")
```

To learn more about these and other customizations, see [the kableExtra vingette](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#Overview).