---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(knitr)
```

### Tidy data principles, reshaping data, and `tidyr`

Packages in the `tidyverse`, like `ggplot2` (useful for visualizing data) and `dplyr` (a key tool in data wrangling), are built to work with so-called "tidy" data. 

```{r, echo = FALSE}
url <- "https://github.com/allisonhorst/stats-illustrations/blob/master/rstats-artwork/tidydata_1.jpg?raw=true"

knitr::include_graphics(url)
```

One of the core arguments for tidy data is often summarized with an adaptation of the famous Leo Tolstoy quote "Happy families are all alike; every unhappy family is unhappy in its own way." 

```{r, echo = FALSE}
url <- "https://github.com/allisonhorst/stats-illustrations/blob/master/rstats-artwork/tidydata_2.jpg?raw=true"

knitr::include_graphics(url)
```

With a consistent set of rules for formatting data, the ideas and tools used to conduct one data analysis can be more easily transferred to another.

```{r, echo = FALSE}
url <- "https://github.com/allisonhorst/stats-illustrations/blob/master/rstats-artwork/tidydata_3.jpg?raw=true"

knitr::include_graphics(url)
```

The `tidyr` package will prove useful in data restructuring or "tidying". All of the packages in the `tidyverse`, including `ggplot2` and `dplyr`, as well as many community-contributed packages, are designed to work  with data in tidy format.

You can see what tidy and untidy data look might look like in practice in the [Tidy example](INTERNAL-LINK) section, and learn how to turn untidy data into tidy data in the [Restructuring with `pivot_wider()` and `pivot_longer()`](INTERNAL-LINK) section.