---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(knitr)
```

### Transforming data with `dplyr`

Once data is tidy, you can use `dplyr` to further transform your data.

The below documentation covers six fundamental `dpylr` functions that should allow you to perform the majority of data wrangling tasks. You can learn about:

  - [Subsetting data with `filter()` and `select()`](INTERNAL-LINK)
  - [Creating new columns with `mutate()`](INTERNAL-LINK)
  - [Summarizing data with `group_by()` and `summarize()`](INTERNAL-LINK)
  - [Arranging rows with `arrange()`](INTERNAL-LINK)
  - [Frequency tables with `count()`](INTERNAL-LINK)
  
Throughout these examples, using the `palmerpenguins` data, you will see the `head()` function used as a way to confirm the effect of the data transformations.

You can also use `head()` to get a sense of your dataset, variables, and values:
```{r}
head(penguins)
```

If you want to learn more about `dplyr`, [the package's website](https://dplyr.tidyverse.org/) contains a more extensive overview of the package and its functions. You might also check out the [data transformation chapter](https://r4ds.had.co.nz/transform.html) of the R for Data Science book.