---
output: html_document
---

```{r, include = FALSE}

```

### From a .csv (with `readr`)

Comma-separated value (`.csv`) files are a very common format of data that can be easily transferred between many different applications, including R. To load a `.csv` file into R, you can use the `readr` package.

First, install the `readr` package:

```{r, eval = FALSE}
install.packages("readr")
```

Now that you have the package installed, load the package into your environment with `library()`.

```{r, warning = FALSE, message = FALSE}
library(readr)
```

Now that you have loaded the package, you can load your data.

First, you will need to know where the data is on your computer. You will need the filepath to the data to show `readr` where to find the file. Since this is a `.csv` file, you will use the `read_csv()` function to read in the file.

Here is an example:

```{r, eval = FALSE}
cat_data <- read_csv("Desktop/Reed/Bio123/cats.csv")
```

The above code reads in the file `cats.csv`, located in a folder `Bio123` within a `Reed` folder on the `Desktop`. The filepath `Desktop/Reed/Bio123` tells `read_csv()` where to find `cats.csv`. 

To find a filepath on computer running MacOSX you can follow these steps:
  - Locate the desired file
  - Right-click on the file
  - Hold down the option key and click "copy 'filename' as Pathname"
  
To find a filepath on a computer running Windows follow these steps:
  - Locate the desired file
  - Hold shift and right-click on the file
  - Select "Copy as Path"
  
The copied value can now be used as the argument in the `read_csv()` function.

`read_csv()` not only finds this data, but also makes it into a dataset, which is then assigned to the name (`cat_data`) using the `<-` operator. This data is now available for use in R (you can confirm this using `View()`).

If you are using the Reed RStudio Server, you can upload your data file to the server ("Files" pane, lower right) and read it in from the server ("Environment" pane, upper left). This process uses the same `read_csv()` function within the server.

If you want to learn more about the `readr` package or the `read_csv()` function, [the Tidyverse documentation for `readr`](https://readr.tidyverse.org/) provides a more extensive overview of the package and its functions.