---
output: html_document
---

```{r, include = FALSE}

```

### From a .csv (with `readr`)

Comma-separated value (`.csv`) files are a very common format of data that can be easily transferred between many different applications, including R. In order to load a `.csv` file into R, you will need the `readr` package.

First, install the `readr` package:

```{r, eval = FALSE}
install.packages("readr")
```

Now that you have the package installed, load the package into your environment with `library()`.

```{r, warning = FALSE, message = FALSE}
library(readr)
```

Now that you have installed the package, you can load your data.

First, you will need to know where the data is on your computer. You will need the filepath to the data to show `readr` where to find the file. Since this is a `.csv` file, you will use the `read_csv()` function to read in the file.

Here is an example:

```{r, eval = FALSE}
cat_data <- read_csv("Desktop/Reed/Bio123/cats.csv")
```

The above code reads in the file `cats.csv`, located in a folder `Bio123` within a `Reed` folder on the `Desktop`. The filepath `Desktop/Reed/Bio123` gives directions to `read_csv()` on where to find `cats.csv`. 

`read_csv()` not only finds this data, but also makes it into a dataset, which is then assigned to the name (`cat_data`) using the `<-` operator. This data is now available for use in R (you can confirm this using `View()`).

If you are using the Reed RStudio Server, you can upload your data file to the server and read it in from the server. This process uses the same `read_csv()` function within the server.

If you want to learn more about the `readr` package or the `read_csv()` function, [the Tidyverse documentation for `readr`](https://readr.tidyverse.org/) provides a more extensive overview of the package and its functions.