---
output: html_document
---

```{r, include = FALSE}

```

### From Excel (with `readxl`)

If you are working with an Excel file, you can read your data into R using the `readxl` package.

```{r, eval = FALSE}
install.packages("readxl")
```

Once the package is installed, load it into your environment with `library()`. 

```{r, warning = FALSE, message = FALSE}
library(readxl)
```

The package `readxl` contains a function called `read_excel()`, which is very similar to [`read_csv()`](link to read_csv doc). The main difference between these two tools is that the file is be a `.xls` or `.xlsx`, rather than `.csv`:

```{r, eval = FALSE}
cat_data <- read_excel("Desktop/Reed/Bio123/cats.xls")
```

If you have multiple sheets in your Excel file, `read_excel()` can access all of them. Let's say the `cats.xls` data has three sheets: the first sheet is about indoor cats, the second is about outdoor cats, and the third is about stray cats. We can load these in as follows:

```{r, eval = FALSE}
indoor_cat <- read_excel(path = "Desktop/Reed/Bio123/cats.xls", sheet = 1)
outdoor_cat <- read_excel(path = "Desktop/Reed/Bio123/cats.xls", sheet = 2)
stray_cat <- read_excel(path = "Desktop/Reed/Bio123/cats.xls", sheet = 3)
```

We can also call the sheets by their names instead of by numbers:

```{r, eval = FALSE}
indoor_cat <- read_excel(path = "Desktop/Reed/Bio123/cats.xls", sheet = "Indoor")
outdoor_cat <- read_excel(path = "Desktop/Reed/Bio123/cats.xls", sheet = "Outdoor")
stray_cat <- read_excel(path = "Desktop/Reed/Bio123/cats.xls", sheet = "Stray")
```

This code will help you to bring in data directly from Excel sheets. If you want to learn more about the `readxl` package or the `read_excel()` function, [the Tidyverse documentation for `readxl`](https://readxl.tidyverse.org/) provides a more extensive overview of the package and its functions.