---
output: html_document
---

```{r, include = FALSE}

```

### From a package

Many R packages contain their own data. Most of the data used in tutorials on the Data at Reed site comes from a package called `palmerpenguins`. The `palmerpenguins` package includes datasets from a study of three different species of penguins found in Antarctica.

```{r, echo = FALSE}
url <- "https://github.com/allisonhorst/palmerpenguins/raw/master/man/figures/lter_penguins.png"

knitr::include_graphics(url)
```
*Artwork by @allison_horst*

To load data from a package, first install the package with the `install.packages()` function:

```{r, eval = FALSE}
install.packages("palmerpenguins")
```

Now that you have the package installed, you need to load the package into your environment with `library()`. You only need to install the package once, but you will need to load the package every time you open R.

```{r}
library(palmerpenguins)
```

Now `palmerpenguins` is loaded. Where is the data?

When working with a package, there are a couple of options for finding the data contained in that package. (While all of these examples focus on the `palmerpenguins` data, you could use these instructions for any dataset of interest.)

For the most complete information, run `help(package = "palmerpenguins")` in your console. This will show you all of the data and functions included in a package, as well as descriptions of what they are.

If you are already familiar with the data and functions in a package and you just need a list of names to jog your memory, you can make use of RStudio's autocomplete. Type the name of your package followed by two colons, like this 

```{r, eval = FALSE}
palmerpenguins::
```

Your tooltip should show you all of the datasets included in the package.

Once you have found the name of the dataset you want, in our case `penguins`,you can use the `data()` function:

```{r}
data(penguins)
```

Depending on how datasets are included in a package, you might run that command and see nothing, or you may see `<Promise>` up in your Environment in the top-right corner of your RStudio window. It might not seem like your data is loaded. One way to confirm your data is loaded is to run the `View()` function in your console:

```{r, eval = FALSE}
View(penguins)
```

Alternatively, you can use the `head()` function to print the first few rows:

```{r}
head(penguins)
```

Your data is loaded! You're ready to go.