---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(knitr)
```

### Combining data

During data analysis, you might find that you have two separate datasets that offer complementary information to each other. In this case, you might be interested in combining, or _joining_ them.

For example, suppose you also have some data about the islands that these penguins live on that you believe might inform your analyses of penguin characteristics.

```{r, include = FALSE}
palmer_islands <-
  tibble(island = c("Torgersen", "Biscoe", "Dream"),
         mean_temperature_c = c(-3, -6, -1),
         mean_elevation_m = c(17, 8, 10))
```

```{r}
palmer_islands
```

`penguins` and `islands` share a column, `island`, that allows you to combine the two datasets. To join these two datasets, we can use the `full_join()` function from `dplyr`.

```{r}
full_join(x = penguins, y = palmer_islands, by = "island") %>%
  select(species, island)
```

"Full" joins are one of many types of joins, and keep any rows that are included in both `x` and `y` (the first and second datasets, respectively). Since the `palmer_islands` dataset only has one row per island, its rows are repeated as many times as needed so that each row in `penguins` has the entry it needs.

Other join types include inner, left, and right joins. These join types differ in how they match rows between the two different datasets. If `x` and `y` are the first and second datasets, respectively,

* `inner_join()` includes all rows in `x` _and_ `y` that match 
* `left_join()` includes all rows in `x` (and any from `y` that match)
* `right_join()` includes all rows in `y` (and any from `x` that match)
* `full_join()` includes all rows in `x` _or_ `y`, even if they do not match something in the other dataset

To read more about combining data, we recommend the [relational data chapter](https://r4ds.had.co.nz/relational-data.html) of R for Data Science.