---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(knitr)
```

#### Creating new columns with `mutate()`

Another common data wrangling task is to create a new variable using the `mutate()` function. When creating a new variable, you provide a name for the new column and a method for calculating the new value. 

Continuing with the penguins data, the below creates a new column for the mean body mass in kilograms:

```{r}
penguins %>%
  mutate(body_mass_kg = (body_mass_g / 1000))
```

Perhaps you realized that all flipper measurements were 4 mm short of the true length; you could use `mutate()` to adjust the data:

```{r}
penguins %>%
  mutate(flipper_length_mm = flipper_length_mm + 4)
```

You can also combine `mutate()` with other functions. The below code calculates total body mass of _all_ penguins on each island.

```{r}
penguins %>%
  group_by(island) %>%
  mutate(island_penguin_mass = sum(body_mass_g))
```

It may be useful to give R rules for creating new variables. For example, the below code divides all penguins into flipper length categories, based on the mean flipper length of the dataset (201mm), using the `case_when()` function.

```{r}
penguins %>%
  mutate(flipper_category =
           case_when( flipper_length_mm > 201 ~ "short",
                      flipper_length_mm < 201 ~ "long",
                      flipper_length_mm == 201 ~ "average"))
```