---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(knitr)
```

#### Creating new columns with `mutate()`

Another common data wrangling task is to create a new variable using the `mutate()` function. When creating a new variable, you provide a name for the new column and a method for calculating the new value. 

Continuing with the penguins data from `palmerpenguins`, the code below creates a new column for the mean body mass in kilograms:

```{r}
penguins %>%
  mutate(body_mass_kg = body_mass_g / 1000)
```

The syntax for mutating a column follows the pattern of `mutate(new_column_name = expression)`, where `expression` is some sort of instruction for combining values in existing columns. In the above example, `new_column_name` is `body_mass_kg`, and `expression` is `body_mass_g / 1000`.


Perhaps you realized that all flipper measurements were 4 mm short of the true length; you could use `mutate()` to adjust the data:

```{r}
penguins %>%
  mutate(flipper_length_mm = flipper_length_mm + 4)
```

You can also combine `mutate()` with other functions. The below code calculates total body mass of _all_ penguins on each island.

```{r}
penguins %>%
  group_by(island) %>%
  mutate(island_penguin_mass = sum(body_mass_g, na.rm = T))
```

It may be useful to give R rules for creating new variables. For example, the below code divides all penguins into flipper length categories, based on the mean flipper length of the dataset (201mm), using the `case_when()` function. You can think of `case_when` as being a multilevel `if` statement. Essentially, the `case_when()` function in the code below is saying "for each observation (row), when the variable `flipper_length_mm` meets a certain condition (greater than, equal to, or less than 201mm), the new column should contain the respective category---`"long"`, `"average"`, or `"short"`.

```{r}
penguins %>%
  mutate(flipper_category =
           case_when( flipper_length_mm > 201 ~ "long",
                      flipper_length_mm == 201 ~ "average",
                      flipper_length_mm < 201 ~ "short"))
```