---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(knitr)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
penguins_sum <- penguins %>%
  group_by(island, year) %>%
  summarize(mean_body_mass_g = mean(body_mass_g, na.rm= TRUE))
```

#### Restructuring with `pivot_wider()` and `pivot_longer()`

One useful data skill is being able to move from untidy data to tidy data and back again. The functions that allow you to do this, `pivot_wider()` and `pivot_longer()`, exist in the package `tidyr`. This package is included in `tidyverse`, along with many other helpful packages. To access these tools, install and load `tidyverse`:

```{r eval = FALSE}
install.packages("tidyverse")
library(tidyverse)
```

With the tools loaded, you can restructure the summary dataset from above. Begin with the first, tidy version:

```{r, message = FALSE, warning = FALSE}
penguins_sum
```

To make the untidy version, "pivot" this data from long to wide format using the `pivot_wider()` function from `tidyr`:

```{r, warning = FALSE, message = FALSE}
penguins_wide <- penguins_sum %>%
  pivot_wider(id_cols = c("island", "year"), 
              names_from = year, 
              values_from = mean_body_mass_g)

penguins_wide
```

Looking more closely at `pivot_wider()`

- `id_cols` (read: ID columns) are the variables that, together, identify what makes a row unique in the original data
- `names_from` determines what variable from the old data will determine the names of columns in the new data
- `values_from` is the name of the variable in the old data that will be contained in cells in the new data

Given this untidy table, you can tidy the data by pivoting from "wide" to "long" using `pivot_longer()`.

```{r}
penguins_wide %>%
  pivot_longer(cols = c("2007", "2008", "2009"), 
               names_to = "year",
               values_to = "mean_body_mass_g")
```

As you might have noticed, `pivot_longer()` and `pivot_wider()` are inverse operations. Pivoting a widened dataset to a longer format gives you back the original dataset, and vice versa.

If you are interested in learning more about tidy data and pivoting, see the [Tidy Data chapter](https://r4ds.had.co.nz/tidy-data.html) in R for Data Science.