---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(knitr)
```

#### Subsetting data with `filter()` and `select()`

Perhaps you wish to subset your data, retaining only observations (rows) that meet given criteria. To subset Rows, use function `filter()` (both words contain an "r", which may help you remember which function to use). The below code subsets only female penguins:

```{r}
penguins %>%
  filter(sex == "female")
```

While this code subsets penguins with a bill length of less than 40mm:

```{r}
penguins %>%
  filter(bill_length_mm < 40)
```

You can also combine requirements for your data ("arguements" in the function), for example subsetting penguins from Torgersen Island that were observed in 2008 with a body mass greater than 3500 grams:

```{r}
penguins %>%
  filter(island == "Torgersen" & year == 2008 & body_mass_g > 3500 )
```

In working with the penguins data, you may have noticed missing values (`NA`) throughout the dataset. Missing data can complicate data analysis, and you should be intentional in how you approach your `NA`s.

You can use `filter()` combined with `!is.na()` ("is not missing") to remove missing observations for a given variable. For example, creating a dataset that is only penguins of known sex:

```{r}
penguins %>%
  filter(!is.na(sex))
```

Conversely, you may wish to subset observations with missing values, using `filter()` in combination with `is.na()` ("is missing").

```{r}
penguins %>%
  filter(is.na(sex))
```

Both logical and Boolean operators can be used with `filter()`. The two pieces of code below (`?base::Logic` and `?Comparison`) give more information on the different operators and their uses. 

```{r, eval = FALSE}
?base::Logic

?Comparison
```

Another way to subset data is by subsetting columns, or variables. A dataset may have more variables than you need, so you can use the `select()` function to subset data to only variables of interest. (Both seleCt and columns contain "c"s, which may help you remember which function to use.)

To subset only species information from your dataset, you can use the following code:

```{r}
penguins %>%
  select(species)
```

You can use a negative selection to subset everything _except_ species information from your data:

```{r}
penguins %>%
  select(-species)
```

You can also list which variables you would like to subset, for example `species`, `island`, and `sex`:

```{r}
penguins %>%
  select(species, island, sex)
```

(To recap, subsetting columns is done with `select` and subsetting rows is done with `filter`; both will be useful as you work with data.)
