---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(knitr)
```

#### Subsetting data with `filter()` and `select()`

Perhaps you wish to subset your data, retaining only observations (rows) that meet given criteria. The below code uses the function `filter()` to subset only female penguins:

```{r}
penguins %>%
  filter(sex == "female")
```

While this code subsets penguins with a bill length of less than 40mm:

```{r}
penguins %>%
  filter(bill_length_mm < 40)
```

You can also get very specific, for example subsetting penguins from Torgersen Island that were observed in 2008 with a body mass greater than 3500 grams:

```{r}
penguins %>%
  filter(island == "Torgersen" & year == 2008 & body_mass_g > 3500 )
```

In working with the penguins data, you may have noticed missing values (`NA`) throughout the dataset. Missing data can complicate data analysis, and you should be intentional in how you approach your `NA`s.

You can use `filter()` combined with `!is.na()` ("is not missing") to remove missing observations for a given variable. For example, creating a dataset that is only penguins of known sex:

```{r}
penguins %>%
  filter(!is.na(sex))
```

Conversely, you may wish to subset observations with missing values, using `filter()` in combination with `is.na()` ("is missing").

```{r}
penguins %>%
  filter(is.na(sex))
```

Another way that you can subset your data is by subsetting columns, or variables. A dataset may have more variables than you need, so you can use the `select()` function to subset data to only variables of interest.

To subset only species information from your dataset, you can use the following code:

```{r}
penguins %>%
  select(species)
```

You can subset everything _except_ species information from your data:

```{r}
penguins %>%
  select(-species)
```

You can also list out which variables you would like to subset, for example `species`, `island`, and `sex`:

```{r}
penguins %>%
  select(species, island, sex)
```


To recap, subsetting columns is done with `select` and subsetting rows is done with `filter`.




