---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(knitr)
```

#### Summarizing data with `group_by()` and `summarize()`

Often, data contains groupings (by geography, time period, experimental group). The `group_by()` function enables R to first group observations and then perform additional calculations. Function `summarize()` often proves useful with `group_by()`, allowing the calculation of measures across groups (mean, median, variance, etc).

Continuing with the penguins data, one goal may be to end up with a variable that shows mean body mass of penguins for each island. First feed the grouping variable into `group_by()`:

```{r}
penguins %>%
  group_by(island)
```

The above code did not change anything about the dataset; the `group_by()` function changes the metadata of the dataset. (You can think of this as R getting the dataset ready for a further transformation to be applied to it, with the rule of "anything that happens, happens within these groupings".) 

With the grouping defined, apply the `summarize()` function to calculate mean body mass by island.

```{r message = FALSE}
penguins %>%
  group_by(island) %>%
  summarise(mean_body_mass_g = mean(body_mass_g, na.rm = TRUE))
```

Missing values (`NA`) are an important consideration in data analysis. In the above code, the `na.rm = TRUE` removes any missing values in `body_mass_g` so R is free to calculate the mean of the body mass data that is in the dataset. (To experiment with the effect of `na.rm = TRUE`, run the above code without that argument.)

You can extend the above approach to group by _both_ island and year:

```{r message = FALSE}
penguins %>%
  group_by(island, year) %>%
  summarise(mean_body_mass_g = mean(body_mass_g, na.rm = TRUE))
```