---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(knitr)
```

### Working with text data with `stringr`

Many of the previous examples have worked with numeric data; you may also work with text, or "string" data. String data can include numbers, letters, spaces, and special characters. Strings are most often represented as `"text surrounded in quotes"` in R. 

The `stringr` package is a useful tool for working with string data. Instead of going through a step-by-step tutorial for each `stringr` function, the below code presents concrete examples using the `penguins` dataset to demonstrate some of the capabilities of `stringr`.

Most functions you will want to use from the `stringr` function begin with `str_`. If you forget the name of the function you need to use, try typing `str_` and scrolling through the auto-completed list of functions exported from `stringr`.

String manipulations are carried out with the use of _patterns_, specified with a language called _regular expressions_ (or "regex", for short). Regular expressions are shorthand ways of specifying general patterns that you wish to match in data. 

Some regex that may be useful for getting started:

1. `^x` : means 'begins with `x`'
2. `x$` : means 'ends with `x`'
3. `x{n}` : means exactly n occurrences of `x`
4. `x{n,}` : means n or more occurrences of `x`
5. `x{n,m}` : between n and m occurrences of `x`
6. `[xyz]` : means 'includes `x` or `y` or `z`'

To see each of these in context, see below.

The below code subsets data (using `filter()`) to penguins of a species which start with the letter 'A', and who are found on an island that ends with the letter 'm'.

```{r}
penguins %>%
  filter(str_detect(species, "^A") & str_detect(island, "m$"))
```

As is often true in coding, there are other ways to arrive at this same result (example: `filter(species == "Adelie" & island == "Dream")`); the above code is meant to provide an extensible example of `str_detect()`.

This same approach can be extended to look for observations that start or end with any given combination of characters. The below will filter to any penguins with a value of `species` beginning with "Adeli".

```{r}
penguins %>%
  filter(str_detect(species, "^Adeli"))
```

The below code filters to species that end in one (1) or or two (2) "o's"; after that subsetting, each species name is capitalized (`Adelie` becomes `ADELIE`) using the `str_to_upper()` function.

```{r}
penguins_edited <- penguins %>%
  filter(str_detect(species, "o{1,2}$"))%>%
  mutate(species = str_to_upper(species))

penguins_edited
```

(If you have reason to have your data in the style of bell hooks, there is also a corresponding `str_to_lower()` function.)

In addition to looking for a pattern (`str_detect()`) or adjusting the case of characters (`str_to_upper()`/`str_to_lower()`), you can also remove characters based on a pattern-matching rule.

The below code uses `str_remove()` to drop two "O's" from the end of each value of `species`. (This can be especially useful when you have observations that have a prefix or a suffix that you do not want.)

```{r}
penguins_edited %>%
  mutate(species = str_remove(species,"O{2}$"))
```

(Note that _no_ values of `species` end in O's, so the above code changes nothing. If a value of `species` ended in, for example, three "O's", the result would only retain one.)

Another common task when working with string data is to replace a pattern, achieved with `str_replace()`.

The below code replaces the "e" at the end of "Biscoe" with "otti".

```{r echo = FALSE}
penguin <- penguins %>%
  select(-body_mass_g)
```

```{r}
penguin %>%
  filter(island == "Biscoe") %>%
  mutate(island = str_replace(island, "e", "tti"))
```

Transforming seabirds into biscuits notwithstanding, knowing how to make replacements such as these will likely be important for your data work.

To replace patterns throughout a dataset, use `str_replace_all()`. The below code replaces all occurrences of vowels with "x".

```{r}
penguins %>%
  mutate(species = str_replace_all(species, "[aeiou]", "x"))
```

The `stringr` package contains many tools beyond those profiled above, and regex can be challenging to navigate. Some additional resources for this work can be found at [stringr cheatsheet](https://evoldyn.gitlab.io/evomics-2018/ref-sheets/R_strings.pdf) (credit: Lise Vaudor). To learn more about `stringr`, [the package's website](https://stringr.tidyverse.org/) contains a more extensive overview of the package and its functions. For further background, visit the [Strings chapter](https://r4ds.had.co.nz/strings.html) of the R for Data Science book.