---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(knitr)
```

#### Tidying example

The "tidy" format of a dataset can depend on how you plan to use the data. For examples of tidy data in action, return to the `penguins` dataset. (For instructions on loading `penguins`, see [Getting Started](LINK-HERE).)

This summarized version of the penguins data, called `penguins_sum`, gives the mean body mass for all penguins on a given island in a given year.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
penguins_sum <- penguins %>%
  group_by(island, year) %>%
  summarize(mean_body_mass_g = mean(body_mass_g, na.rm= TRUE))

penguins_sum
```

The resulting data is "tidy": each variable has its own column, each observation (penguin) is a row, and each measurement has its own cell. 

Compare that with this representation of the same data:

```{r, echo = FALSE}
penguins_sum %>%
  pivot_wider(id_cols = c("island", "year"), names_from = year, values_from = mean_body_mass_g)
```

While the above data contains the same information, in this version, the mean body mass is found in three different columns, one for each year. This representation is usually called "wide" data, while the first option is referred to as "long." Per tidy data rules, `year` and `body mass` are both variables and should therefore have their own columns. Similarly, `year` should have its own column. `tidyr` will help us "pivot" from one to the other.