---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(knitr)
```

# Wrangling Data

### Tidy data principles, reshaping data, and `tidyr`

Packages in the `tidyverse`, like `ggplot2` (useful for visualizing data) and `dplyr` (a key tool in data wrangling), are built to work with so-called "tidy" data. 

```{r, echo = FALSE}
url <- "https://github.com/allisonhorst/stats-illustrations/blob/master/rstats-artwork/tidydata_1.jpg?raw=true"

knitr::include_graphics(url)
```

One of the core arguments for tidy data is often summarized with an adaptation of the Leo Tolstoy quote "Happy families are all alike; every unhappy family is unhappy in its own way." 

```{r, echo = FALSE}
url <- "https://github.com/allisonhorst/stats-illustrations/blob/master/rstats-artwork/tidydata_2.jpg?raw=true"

knitr::include_graphics(url)
```

With a consistent set of rules for formatting data, the ideas and tools used to conduct one data analysis can be more easily transferred to another.

```{r, echo = FALSE}
url <- "https://github.com/allisonhorst/stats-illustrations/blob/master/rstats-artwork/tidydata_3.jpg?raw=true"

knitr::include_graphics(url)
```

The `tidyr` package will prove useful in data restructuring or "tidying". All of the packages in the `tidyverse`, including `ggplot2` and `dplyr`, as well as many community-contributed packages, are designed to work with data in tidy format.

---

#### Tidying example

The "tidy" format of a data can depend on your intended use of the data. To get started in thinking about tidy data, return to the `penguins` data. (For instructions on loading `penguins`, see [Getting Started](LINK-HERE).)

This summarized version of the penguins data, called `penguins_sum`, gives the mean body mass for all penguins on a given island in a given year.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
penguins_sum <- penguins %>%
  group_by(island, year) %>%
  summarize(mean_body_mass_g = mean(body_mass_g, na.rm= TRUE))

penguins_sum
```

The resulting data is tidy: each variable has its own column, each observation (penguin) is a row, and each measurement has its own cell. 

Compare that with this representation of the same data:

```{r, echo = FALSE}
penguins_sum %>%
  pivot_wider(id_cols = c("island", "year"), names_from = year, values_from = mean_body_mass_g)
```

While the above example contains the same information, in this version the mean body mass is found in three different columns, one for each year. This representation is usually called "wide" data, while the first option is referred to as "long." Per tidy data rules, `year` and `body mass` are both variables and should therefore have their own columns. Similarly, `year` should have its own column. `tidyr` will help us "pivot" from one to the other.

---

#### Restructuring with `pivot_wider()` and `pivot_longer()`

Let's take a look at how to move from untidy data to tidy data and back again. The functions that allow you to do this, `pivot_wider()` and `pivot_longer()`, exist in the package `tidyr`. This package is included in `tidyverse`, along with many other helpful packages, so you can install and load `tidyverse`:

```{r eval = FALSE}
install.packages("tidyverse")
library(tidyverse)
```

To restructure the data, begin with the first, tidy version:

```{r, message = FALSE, warning = FALSE}
penguins_sum
```

To make the untidy version, "pivot" this data from long to wide format using the `pivot_wider()` function from `tidyr`:

```{r, warning = FALSE, message = FALSE}
penguins_wide <- penguins_sum %>%
  pivot_wider(id_cols = c("island", "year"), 
              names_from = year, 
              values_from = mean_body_mass_g)

penguins_wide
```

Looking more closely at `pivot_wider()`

- `id_cols` (read: "ID columns") are the variables that identify what makes a row unique in the original data
- `names_from` determines what variable from the old data will determine the names of columns in the new data
- `values_from` is the name of the variable in the old data that will be contained in cells in the new data

Given this untidy table, you can tidy the data by pivoting from "wide" to "long" using `pivot_longer()`.

```{r}
penguins_wide %>%
  pivot_longer(cols = c("2007", "2008", "2009"), 
               names_to = "year",
               values_to = "mean_body_mass_g")
```

As you might have noticed, `pivot_longer()` and `pivot_wider()` are inverse operations. Pivoting a widened dataset to a longer format gives you back the original dataset, and vice versa.

If you are interested in learning more about tidy data and pivoting, see the [Tidy Data chapter](https://r4ds.had.co.nz/tidy-data.html) in R for Data Science.

---

### Transforming data with `dplyr`

Once data is tidy, you can use `dplyr` to further transform your data.

The below documentation covers six fundamental `dpylr` functions that should allow you to perform the majority of data wrangling tasks. Throughout these examples, using the `palmerpenguins` data, you will see the `head()` function used as a way to confirm the effect of the data transformations.

You can also use `head()` to get a sense of your dataset, variables, and values:
```{r}
head(penguins)
```

If you want to learn more about `dplyr`, [the package's website](https://dplyr.tidyverse.org/) contains a more extensive overview of the package and its functions. You might also check out the [data transformation chapter](https://r4ds.had.co.nz/transform.html) of the R for Data Science book.

---

#### Subsetting data with `filter()` and `select()`

Perhaps you wish to subset your data, retaining only observations (rows) that meet given criteria. The below code uses the function `filter()` to subset only female penguins:

```{r}
penguins %>%
  filter(sex == "female")
```

While this code subsets penguins with a bill length of less than 40mm:

```{r}
penguins %>%
  filter(bill_length_mm < 40)
```

You can also get very specific, for example subsetting penguins from Torgersen Island that were observed in 2008 with a body mass greater than 3500 grams:

```{r}
penguins %>%
  filter(island == "Torgersen" & year == 2008 & body_mass_g > 3500 )
```

In working with the penguins data, you may have noticed missing values (`NA`) throughout the dataset. Missing data can complicate data analysis, and you should be intentional in how you approach your `NA`s.

You can use `filter()` combined with `!is.na()` ("is not missing") to remove missing observations for a given variable. For example, creating a dataset that is only penguins of known sex:

```{r}
penguins %>%
  filter(!is.na(sex))
```

Conversely, you may wish to subset observations with missing values, using `filter()` in combination with `is.na()` ("is missing").

```{r}
penguins %>%
  filter(is.na(sex))
```

Another way that you can subset your data is by focusing on columns. A dataset may have more variables than you need; if that is the case, use the `select()` function to subset data to only variables (columns) of interest.

For example, to subset only species information from your dataset, use the following code:

```{r}
penguins %>%
  select(species)
```

You can also use `select()` in the negative, subsetting everything _except_ species information from your data:

```{r}
penguins %>%
  select(-species)
```

You can subset a number of variables, for example `species`, `island`, and `sex`. 

```{r}
penguins %>%
  select(species, island, sex)
```

R will order the resulting data in the order you list the variables.

---

#### Creating new columns with `mutate()`

Another common data wrangling task is to create a new variable using the `mutate()` function. When creating a new variable, provide a name for the new column and a method for calculating the new value. 

Continuing with the penguins data, the below creates a new column for the mean body mass in kilograms:

```{r}
penguins %>%
  mutate(body_mass_kg = (body_mass_g / 1000))
```

Perhaps you realized that all flipper measurements were 4 mm short of the true length; you could use `mutate()` to adjust the data:

```{r}
penguins %>%
  mutate(flipper_length_mm = flipper_length_mm + 4)
```

You can also combine `mutate()` with other functions. The below code calculates total body mass of _all_ penguins on each island.

```{r}
penguins %>%
  group_by(island) %>%
  mutate(island_penguin_mass = sum(body_mass_g))
```

It may be useful to give R rules for creating new variables. For example, the below code divides all penguins into flipper length categories, based on the mean flipper length of the dataset (201mm), using the `case_when()` function.

```{r}
penguins %>%
  mutate(flipper_category =
           case_when( flipper_length_mm > 201 ~ "short",
                      flipper_length_mm < 201 ~ "long",
                      flipper_length_mm == 201 ~ "average"))
```

---

#### Summarizing data with `group_by()` and `summarize()`

Often, data contains groupings (e.g. by geography, time period, experimental group). The `group_by()` function tells R to first group observations and then perform additional calculations. Function `summarize()` often proves useful with `group_by()`, allowing the calculation of measures (mean, median, variance, etc) across groups.

Continuing with the penguins data, you may wish to calculate mean body mass of penguins for each island. First feed the grouping variable into `group_by()`:

```{r}
penguins %>%
  group_by(island)
```

The above code did not change anything about the dataset; the `group_by()` function changes the metadata of the dataset. (You can think of this as R getting the dataset ready for a further transformation to be applied to it, with the rule of "anything that happens, happens within these groupings".) 

With the grouping defined, apply the `summarize()` function to calculate mean body mass by island.

```{r message = FALSE}
penguins %>%
  group_by(island) %>%
  summarize(mean_body_mass_g = mean(body_mass_g, na.rm = TRUE))
```

Missing values (`NA`) are an important consideration in data analysis. In the above code, the `na.rm = TRUE` removes any missing values in `body_mass_g` so R is free to calculate the mean of the body mass data that is in the dataset. (To experiment with the effect of `na.rm = TRUE`, run the above code without that argument.)

You can extend the above approach to group by _both_ island and year:

```{r message = FALSE}
penguins %>%
  group_by(island, year) %>%
  summarize(mean_body_mass_g = mean(body_mass_g, na.rm = TRUE))
```

---

#### Arranging rows with `arrange()`

Sometimes it is helpful to sort data, alphabetically or numerically. You can sort data using the `arrange()` function from `dplyr`.

To arrange all penguins, in order of increasing bill length:

```{r}
penguins %>%
  arrange(bill_length_mm)
```

In order or decreasing bill length:

```{r}
penguins %>%
  arrange(desc(bill_length_mm))
```

For text data, `dplyr` will arrange alphabetically.

```{r}
penguins %>%
  arrange(species)
```

---

#### Frequency tables with `count()`

Frequency tables provide one way of displaying information about your data. For example, how many male and female penguins are in the sample?

First, use function `count()` by the variable of interest (`sex`):

```{r}
penguins %>%
  count(sex)
```

The penguins dataset includes 165 females, 168 males, and 11 individual birds for which sex is unknown.

To divide these counts by island (as well as sex), `count()` by `island` and `sex`:

```{r}
penguins %>%
  count(island, sex)
```

`count()` is a quick way to produce frequency tables in R.

---

### Combining data

During data analysis, you might find that you have two separate datasets that offer complementary information to each other. In this case, you might be interested in combining, or _joining_ them.

For example, suppose you also have some data about the islands that these penguins live on that you believe might inform your analyses of penguin characteristics.

```{r, include = FALSE}
palmer_islands <-
  tibble(island = c("Torgersen", "Biscoe", "Dream"),
         mean_temperature_c = c(-3, -6, -1),
         mean_elevation_m = c(17, 8, 10))
```

```{r}
palmer_islands
```

`penguins` and `islands` share a column, `island`, that allows you to combine the two datasets. To join these two datasets, use the `full_join()` function from `dplyr`.

```{r}
full_join(x = penguins, y = palmer_islands, by = "island") %>%
  select(species, island)
```

"Full" joins are one of many types of joins, and keep any rows that are included in both `x` and `y` (the first and second datasets, respectively). Since the `palmer_islands` dataset only has one row per island, its rows are repeated as many times as needed so that each row in `penguins` has the entry it needs.

Other join types include inner, left, and right joins. These join types differ in how they match rows between the two different datasets. If `x` and `y` are the first and second datasets, respectively,

* `inner_join()` includes all rows in `x` _and_ `y` that match 
* `left_join()` includes all rows in `x` (and any from `y` that match)
* `right_join()` includes all rows in `y` (and any from `x` that match)
* `full_join()` includes all rows in `x` _or_ `y`, even if they do not match something in the other dataset

To read more about combining data, see the [relational data chapter](https://r4ds.had.co.nz/relational-data.html) of R for Data Science.

---

### Working with text data with `stringr`

Many of the previous examples have worked with numeric data. You may also work with text, or "string" data. String data can include numbers, letters, spaces, and special characters. Strings are most often represented as `"text surrounded in quotes"` in R. 

The `stringr` package is a useful tool for working with string data. Instead of going through a step-by-step tutorial for each `stringr` function, the below code presents concrete examples using the `penguins` dataset to demonstrate some of the capabilities of `stringr`.

Most functions you will want to use from the `stringr` function begin with `str_`. If you forget the name of the function you need to use, try typing `str_` and scrolling through the auto-completed list of functions exported from `stringr`.

String manipulations are carried out with the use of _patterns_, specified with a language called _regular expressions_ (or "regex" for short). Regular expressions specify patterns that you wish to match in data. 

Some regex that may be useful for getting started:

1. `^x` : means 'begins with `x`'
2. `x$` : means 'ends with `x`'
3. `x{n}` : means exactly n occurrences of `x`
4. `x{n,}` : means n or more occurrences of `x`
5. `x{n,m}` : between n and m occurrences of `x`
6. `[xyz]` : means 'includes `x` or `y` or `z`'

To see each of these in context, see below.

The below code subsets data (using `filter()`) to penguins of a species which start with the letter 'A', and who are found on an island that ends with the letter 'm'.

```{r}
penguins %>%
  filter(str_detect(species, "^A") & str_detect(island, "m$"))
```

As is often true in coding, there are other ways to arrive at this same result (example: `filter(species == "Adelie" & island == "Dream")`); the above code is meant to provide an extensible example of `str_detect()`.

This same approach can be extended to look for observations that start or end with any given combination of characters. The below will filter to any penguins with a value of `species` beginning with "Adeli".

```{r}
penguins %>%
  filter(str_detect(species, "^Adeli"))
```

The below code filters to species that end in one (1) or or two (2) "o's"; after that subsetting, each species name is capitalized (`Adelie` becomes `ADELIE`) using the `str_to_upper()` function.

```{r}
penguins_edited <- penguins %>%
  filter(str_detect(species, "o{1,2}$"))%>%
  mutate(species = str_to_upper(species))

penguins_edited
```

(If you have reason to have your data in the style of bell hooks, there is also a corresponding `str_to_lower()` function.)

In addition to looking for a pattern (`str_detect()`) or adjusting the case of characters (`str_to_upper()`/`str_to_lower()`), you can also remove characters based on a pattern-matching rule.

The below code uses `str_remove()` to drop two "O's" from the end of each value of `species`. (This can be especially useful when you have observations that have a prefix or a suffix that you do not want.)

```{r}
penguins_edited %>%
  mutate(species = str_remove(species,"O{2}$"))
```

(Note that _no_ values of `species` end in O's, so the above code changes nothing. If a value of `species` ended in, for example, three "O's", the result would only retain one.)

Another common task when working with string data is to replace a pattern, achieved with `str_replace()`.

The below code replaces the "e" at the end of "Biscoe" with "otti".

```{r echo = FALSE}
penguin <- penguins %>%
  select(-body_mass_g)
```

```{r}
penguin %>%
  filter(island == "Biscoe") %>%
  mutate(island = str_replace(island, "e", "tti"))
```

Transforming islands into biscuits notwithstanding, knowing how to make replacements such as these will likely be important for your data work.

To replace patterns throughout a dataset, use `str_replace_all()`. The below code replaces all occurrences of vowels with "x".

```{r}
penguins %>%
  mutate(species = str_replace_all(species, "[aeiou]", "x"))
```

The `stringr` package contains many tools beyond those profiled above, and regex can be challenging to navigate. Some additional resources for this work can be found at [stringr cheatsheet](https://evoldyn.gitlab.io/evomics-2018/ref-sheets/R_strings.pdf) (credit: Lise Vaudor). To learn more about `stringr`, [the package's website](https://stringr.tidyverse.org/) contains a more extensive overview of the package and its functions. For further background, visit the [Strings chapter](https://r4ds.had.co.nz/strings.html) of the R for Data Science book.
